% The volume-uniform Poincare walls — machine-checked obstruction paper.
% Lane: hRpoly / P4-CT / P4-ADJ / W-1..W-3c (ledger Addenda 467-507).
% REGIME: tricotomy; negative results presented as theorems, not failures;
% NO scores, NO desk language, NO commit chronology beyond the
% reproducibility freeze.  Freeze: Lean tree of source checkpoint
% a17d7816 (sealed 33072662); this manuscript revision added in a strictly
% editorial later commit.
\documentclass[11pt]{article}
\usepackage[margin=1.1in]{geometry}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{booktabs,longtable,array}
\usepackage[colorlinks=true,linkcolor=blue,urlcolor=blue,citecolor=blue]{hyperref}

\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{remark}[theorem]{Remark}
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}

\emergencystretch=3em
\tolerance=2000

\newcommand{\lean}[1]{\texttt{#1}}
\newcommand{\leanprooflink}[2]{%
  \href{https://github.com/lluiseriksson/THE-ERIKSSON-PROGRAMME/blob/a17d78165eaf7fc9ad505e8d3ed8e544848f8518/YangMills/RG/#1.lean}{\nolinkurl{#2}}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\SU}{\mathrm{SU}}
\newcommand{\su}{\mathfrak{su}}
\newcommand{\CP}{C_{\mathrm{P}}}
\newcommand{\inn}[2]{\langle #1,\, #2\rangle}

\title{The Volume-Uniform Poincar\'e Walls:\\
Machine-Checked Obstructions for Flat and Fluctuation-Sector\\
Block-Poincar\'e Routes\\
to Combes--Thomas Coercivity in Lattice Yang--Mills}
\author{Lluis Eriksson}
\date{July 14, 2026 (version 1.2)}

\begin{document}
\maketitle

\begin{abstract}
Inside a Lean~4 formalization programme for four-dimensional $\SU(N_c)$
lattice Yang--Mills, we report two machine-checked \emph{negative}
results and the machine-checked infrastructure that makes them
meaningful.  The positive substrate is: (i) a fixed-volume
Combes--Thomas chain for
self-adjoint coercive finite-range lattice operators, instantiated on the
flat gauge-fixed covariance of the physical shell, with coercivity
constant $c = \min(1,a)/\CP$ fed by a proved fixed-volume flat
Hodge/block-Poincar\'e inequality; and (ii) the concrete adjoint model of
$\SU(n)$ --- $\su(n)$ with the trace inner product,
$\dim_{\R}\su(n)=n^2-1$, and the isometric transport to Euclidean
coordinates --- so that the abstract adjoint-model interface has a
concrete nontrivial inhabitant and the flat-lane results can be
instantiated with the genuine matricial adjoint model.  The first wall
states that, under the block normalization actually used by the
formalized chain, every flat
Hodge/block-Poincar\'e constant obeys $L^d/L^2 \le \CP$ on the fine torus
of side $LN'$, hence the volume-uniform Poincar\'e gate is \emph{provably
false} for $d\ge 3$, $N_c\ge 2$, and no positive coercivity constant
survives all volumes through this route.  The route consumed by the
fixed-volume endpoint is therefore closed, by theorem rather than by
impression.  A second wall stands in the fluctuation sector.  For
$d\ge3$ and a transported half-period square-wave mode on the exact fine
side $(2M)N'$, the formalization proves
$\|QA\|^2\le (2M)^{-1}\|A\|^2$, the exact identity
$\inn{A}{K_0A}=8((2M)N')^{-1}\|A\|^2$, and therefore a Rayleigh numerator
at most $9(2M)^{-1}\|A\|^2$.  Every quotient Poincar\'e constant is thus
at least $2M/9$, so the volume-uniform fluctuation-sector gate is also
provably false for every positive $N'$, $d\ge3$, $N_c\ge2$, and every
adjoint model.  Everything stated here is checked
by Lean~4 against a pinned Mathlib, with zero \lean{sorry}, zero project
axioms, and a committed axiom-oracle transcript; a theorem--artifact map
and reproduction instructions are included.  No claim toward the
continuum Yang--Mills construction or a mass-gap theorem is made; both
remain outside the scope of this work.
\end{abstract}

\section{Introduction}

Constructive approaches to the four-dimensional Yang--Mills mass gap in
the Balaban tradition~\cite{Balaban84a,Balaban84b} route spectral
information through renormalization-group covariances whose kernels must
decay \emph{uniformly in the volume}.  A standard engine for such decay
is the Combes--Thomas argument~\cite{CombesThomas}: a self-adjoint
coercive finite-range operator has an inverse whose kernel entries decay
exponentially in operator norm, at a rate governed by the coercivity
constant.  Whoever supplies
the coercivity therefore controls the physics; and in a gauge-fixed
lattice formalization the natural supplier is a Poincar\'e-type
inequality tying the field norm to a Hodge term plus a block-averaging
constraint.

This paper reports what happened when that plan was executed
\emph{formally}, in Lean~4~\cite{Lean4} against
Mathlib~\cite{mathlib}, inside a long-running formalization programme
whose earlier machine-checked artifacts include a volume-uniform
Wilson-loop area law via a formalized cluster
expansion~\cite{ErikssonVU}, rooted-tree majorants for polymer expansions
with holes~\cite{ErikssonHoles}, a mechanized Bessel-ratio calculus for
lattice gauge expansions~\cite{ErikssonC4,ErikssonC3,ErikssonC2}, an
exact two-dimensional $\SU(2)$ theory~\cite{Eriksson2D}, and the audited
programme map~\cite{ErikssonMap}.

The outcome has four layers, and we state them with the labels they
deserve.

\textbf{Positive (fixed volume).}  A complete Combes--Thomas chain
for coercive finite-range operators on finite metric graphs,
instantiated at the flat gauge-fixed covariance of the physical shell:
for every fixed volume there exist a positive rate $\theta$ and amplitude
$2/c$ with
$\bigl\|(K^{-1})_{q,p}\bigr\|_{\mathrm{op}} \le (2/c)\,e^{-\theta\, d(q,p)}$
for all bond pairs $(q,p)$,
where $c = \min(1,a)/\CP$ and $\CP$ is a \emph{proved} fixed-volume flat
Hodge/block-Poincar\'e constant (Section~\ref{sec:ct}).

\textbf{Positive (non-vacuity).}  The abstract adjoint-model interface
consumed by the chain is inhabited by the \emph{true} matricial adjoint
model of $\SU(n)$: $\su(n)$ as the real inner-product space of traceless
skew-Hermitian matrices under the trace form, its real dimension
$n^2-1$, and the isometric transport to
$\R^{n^2-1}$ carrying $\mathrm{Ad}(g)X = gXg^{\dagger}$
(Section~\ref{sec:adj}).  The flat-lane results also admit the trivial
inhabitant, since only $\mathrm{Ad}(1)$ is consumed there; the matricial
construction removes the intended-model vacuity and prepares the
nontrivial-background lane.

\textbf{Negative (the constant-sector wall).}  Under the block
normalization actually used by the formalized chain --- the unscaled
line-integral block map, which sends constants to $L$ times constants ---
\emph{every} flat Hodge/block-Poincar\'e constant obeys
\[
\frac{L^d}{L^2} \;\le\; \CP
\]
on the fine torus of side $N = LN'$.  Consequently (Section~\ref{sec:wall}):
the volume-uniform Poincar\'e gate is \emph{false} for $d \ge 3$,
$N_c\ge 2$; the Combes--Thomas coercivity fed by this route obeys
$c \le \min(1,a)\,L^2/L^d \to 0$; and no single $c_0>0$ is dominated at
every volume by the route's coercivity.  The per-volume constants
\emph{do} exist, so the obstruction kills a real route, not an empty one.

\textbf{Negative (the fluctuation-sector wall).}  The first witness is
excluded by restricting to cochains orthogonal to every constant
generator (Section~\ref{sec:quotient}); this non-transfer alone was
result-neutral.  The registered one-sided falsifier is then evaluated
completely (Section~\ref{sec:falsifier}).  Its Rayleigh numerator decays
like $(2M)^{-1}$ along even block scales, forcing
$2M/9\le \CP$ and refuting the volume-uniform quotient gate for every
positive fixed coarse side $N'$.  Thus removing the constant sector does
not rescue the current unscaled block normalization.  The argument still
respects the falsifier's asymmetry: divergence refutes the gate, whereas a
bounded quotient on this mode would not have proved it.

\subsection*{Honest scope}

This paper proves no mass gap, no volume-uniform Combes--Thomas bound,
and nothing about the continuum limit; no claim toward the continuum
Yang--Mills construction or a mass-gap theorem~\cite{JaffeWitten} is
made --- both remain outside the scope of this work.
Both walls are theorems about \emph{one specific normalization} --- the
current unscaled line-integral block map with the current unweighted
coarse norm.  The second wall shows that merely restricting this same
form to the fluctuation sector does not rescue it.  The results do not
exclude a differently rescaled or weighted block operator, whose gate
must be formulated and audited anew, or coercivity extracted directly
from the interacting Wilson Hessian.  Their value is navigational: they
convert two would-be silent dead ends into machine-checked boundaries,
before any downstream construction is built on them.

\section{The formal setting}\label{sec:setting}

All objects below are Lean definitions in the repository accompanying
this paper (Section~\ref{sec:repro}); we give their mathematical content
and their Lean names.

\paragraph{Cochains.}  Fix $d,N,N_c\in\mathbb{N}$, all positive.  Sites
form the discrete torus $\Lambda_N=(\Z/N)^d$
(\lean{FinBox d N}); positive bonds are pairs $b=(x,k)$ of a site and a
direction (\lean{PhysicalBond d N}).  A \emph{physical one-cochain} is a
map $A$ from positive bonds to the internal coordinate space
$\R^{N_c^2-1}$ (\lean{SUNLieCoord Nc}), with the $\ell^2$ inner product
in both the bond and internal indices
(\lean{PhysicalGaugeOneCochain d N Nc}).

\paragraph{The block map.}  For $N=LN'$, the flat block constraint
$Q\colon$ fine cochains (side $LN'$) $\to$ coarse cochains (side $N'$)
assigns to a coarse bond the \emph{unscaled} line integral (sum) of the
$L$ fine bond values along the corresponding block line
(\lean{flatBlockConstraintQCLM}, built on \lean{linAvg}).  The
normalization fact that drives everything below is exact
(\lean{flatBlockConstraintQCLM\_constant}):
\begin{equation}\label{eq:Qconst}
Q\,(A_v) \;=\; L\cdot A^{\mathrm{coarse}}_v,
\end{equation}
where $A_v$ denotes the direction-wise constant cochain
$A_v(x,k) = v_k$ for a fixed family $v\colon \{1,\dots,d\}\to\R^{N_c^2-1}$
(\lean{constantPhysicalGaugeOneCochain}).

\paragraph{The flat Hodge operator.}  At the trivial gauge background,
$K_0 = D_1^{\dagger}D_1 + D_0 D_0^{\dagger}$
(\lean{flatGaugeHodgeK0CLM}), so that
$\inn{A}{K_0 A} = \|\mathrm{curl}\,A\|^2 + \|\mathrm{div}\,A\|^2$ with
the standard lattice curl on plaquettes and backward divergence on sites
(\lean{flatGaugeHodgeK0\_inner\_right}).  Constants are flat-harmonic:
$K_0 A_v = 0$
(\lean{flatGaugeHodgeK0CLM\_constantPhysicalGaugeOneCochain}).

\paragraph{The Poincar\'e predicate.}  For an adjoint model $\rho$
(Section~\ref{sec:adj}) and $\CP\in\R$,
\lean{FlatGaugeHodgePoincare d L N' Nc $\rho$ CP} is the statement
\begin{equation}\label{eq:poincare}
0 < \CP \quad\wedge\quad
\forall A:\ \|A\|^2 \;\le\; \CP\,\bigl(\inn{A}{K_0A} + \|QA\|^2\bigr),
\end{equation}
over fine cochains at side $LN'$.  The repository proves, at every fixed
volume, that such constants exist
(\lean{exists\_flatGaugeHodgePoincare}); this non-vacuity is what makes
the negative results of Section~\ref{sec:wall} statements about a live
route.

\section{The fixed-volume Combes--Thomas endpoint}\label{sec:ct}

The Combes--Thomas chain is formalized for abstract self-adjoint
coercive finite-range operators on finite metric graphs: the tilt
algebra and tilted entry bounds, the block Schur bound, coercivity
survival under tilting at the budget
$M(e^{\theta R}-1)N_R \le c/2$, the tilted inverse with
$\|K_\theta^{-1}\|\le 2/c$, and the kernel extraction at a root
(\lean{PhysicalCoerciveCombesThomas*.lean}).  Instantiated at the flat
gauge-fixed covariance of the physical shell (stencils and locality
proved, not assumed), the endpoint reads:

\begin{theorem}[fixed-volume Combes--Thomas;
\lean{flatGaugeFixedCovariance\_CT\_fixedVolume},
\lean{exists\_flatGaugeFixedCovariance\_CT\_fixedVolume}]
Fix the volume parameters $(d,L,N')$, a mass $a>0$, and a flat
Hodge/block-Poincar\'e constant $\CP$ as in \eqref{eq:poincare}.  Let
$c=\min(1,a)/\CP$.  Then there exists $\theta>0$ meeting the explicit
tilt budget such that for all bond pairs $(q,p)$ and every internal
vector $v$,
\[
\bigl\|(K^{-1})_{q,p}\, v\bigr\| \;\le\; \frac{2}{c}\;
e^{-\theta\, d(q,p)}\,\|v\|,
\]
equivalently
$\|(K^{-1})_{q,p}\|_{\mathrm{op}} \le (2/c)\,e^{-\theta\,d(q,p)}$ ---
the form quantified literally by
\lean{PhysicalCovarianceExponentialKernelBound}.
\end{theorem}

Fixed volume means fixed volume: $\CP$, hence $c$ and $\theta$, may
depend on $(d,L,N',a)$.  The question the rest of this paper answers is
whether \emph{this} $c$ can be made volume-uniform.  It cannot.

\section{The true adjoint model of $\SU(n)$}\label{sec:adj}

The chain consumes an abstract interface \lean{SUNAdjointModel n}: a
family of isometries $\mathrm{Ad}(g)$ of $\R^{n^2-1}$ indexed by
$g\in\SU(n)$ with $\mathrm{Ad}(1)=\mathrm{id}$,
$\mathrm{Ad}(gh)=\mathrm{Ad}(g)\mathrm{Ad}(h)$, and inner-product
invariance.  A trivial inhabitant ($\mathrm{Ad}(g)=\mathrm{id}$) exists
and suffices for the flat lane, where only $\mathrm{Ad}(1)$ is consumed;
leaving matters there would invite a vacuity objection the moment a
non-trivial background appears.  The repository therefore constructs the
genuine object:

\begin{theorem}[the matricial adjoint model;
\lean{matrixSUNAdjointModel}]\label{thm:adj}
Let $\su(n)$ be the real vector space of traceless skew-Hermitian
$n\times n$ complex matrices with the trace form
$\inn{X}{Y} = \operatorname{Re}\operatorname{tr}(X^{\dagger}Y)$
(\lean{SUNAdjointMatrixSubstrate}, \lean{SUNAdjointInnerSpace}).  Then
$\dim_{\R}\su(n) = n^2-1$ (\lean{finrank\_suLie}), the reindexed
orthonormal-basis representation gives a linear isometric equivalence
$\su(n)\simeq\R^{n^2-1}$ (\lean{suLieCoordIso}), and conjugating the
adjoint action $X\mapsto gXg^{\dagger}$ by this isometry yields a
\lean{SUNAdjointModel n} whose action laws and inner-product invariance
are proved, not hypothesized.
\end{theorem}

The dimension count is itself formalized (Hermitian/skew decomposition,
multiplication-by-$i$ equivalence, ambient real dimension $2n^2$, and
rank--nullity through the imaginary-trace functional with explicit
surjectivity witness $i\cdot\mathbf{1}$).  The coordinate choice is the
non-canonical standard orthonormal basis; any other choice conjugates
the model by a fixed orthogonal map, and nothing downstream depends on
it.

\section{The wall}\label{sec:wall}

Everything in this section is proved in
\lean{PhysicalGaugeFlatPoincare.lean} (the fixed-volume audit) and
\lean{PhysicalPoincareWall.lean} (the volume-uniform obstruction).

\subsection{The constant sector forces the constant}

\begin{lemma}[exact sector normalization;
\lean{flatBlockConstraint\_controls\_constantSector}]\label{lem:sector}
For every $v$,
\[
\|A_v\|^2 \;=\; \frac{L^d}{L^2}\,\|Q A_v\|^2 .
\]
\end{lemma}

\begin{proof}
$\|A_v\|^2 = N^d\sum_k\|v_k\|^2$ with $N=LN'$
(\lean{norm\_sq\_constantPhysicalGaugeOneCochain}), while by
\eqref{eq:Qconst},
$\|QA_v\|^2 = L^2 N'^d \sum_k\|v_k\|^2$
(\lean{flatBlockConstraintQCLM\_constant\_norm\_sq}); divide.
\end{proof}

\begin{theorem}[the wall, fixed volume;
\lean{flatPoincare\_constantSector\_lower\_bound}]\label{thm:wall}
Let $N_c\ge 2$ and let $\CP$ satisfy \eqref{eq:poincare}.  Then
\[
\frac{L^d}{L^2}\;\le\;\CP .
\]
\end{theorem}

\begin{proof}
For $N_c \ge 2$ the internal space $\R^{N_c^2-1}$ is nontrivial, so
there is $v$ with $\sum_k \|v_k\|^2>0$
(\lean{exists\_pos\_constantSector}; for $N_c=1$, $\su(1)=0$ and the
sector is empty --- the hypothesis is stated, not hidden).  Apply
\eqref{eq:poincare} to $A_v$: the Hodge term vanishes ($K_0A_v=0$), so
$\|A_v\|^2 \le \CP\|QA_v\|^2$, and Lemma~\ref{lem:sector} gives the
claim after cancelling $\|QA_v\|^2>0$.
\end{proof}

\begin{corollary}[linear form;
\lean{flatPoincare\_linear\_lower\_bound}]\label{cor:linear}
If moreover $d\ge 3$, then $L \le \CP$.  (For $d=4$ the forced growth is
$L^2$; the linear form is all the divergence the obstruction needs.
$d=2$ is the scale-invariant exemption $L^d/L^2=1$ and is deliberately
not claimed.)
\end{corollary}

\subsection{The route's coercivity dies with the volume}

\begin{theorem}[coercivity upper bound;
\lean{flatPoincare\_coercivity\_upper\_bound}]
For $a>0$, $N_c\ge2$, and any $\CP$ as in \eqref{eq:poincare},
\[
\frac{\min(1,a)}{\CP} \;\le\; \min(1,a)\,\frac{L^2}{L^d}.
\]
\end{theorem}

\begin{theorem}[no volume-uniform coercivity through flat Poincar\'e]
\label{thm:noc}
{\footnotesize\emph{Lean:}
\path{no_volumeUniform_coercivity_via_flatPoincare}.}\par
Let $d\ge3$, $N_c\ge2$, $a>0$.  Suppose $c_0\in\R$ is dominated at every
volume: for every $L\in\mathbb{N}$ there exists $\CP(L)$ satisfying
\eqref{eq:poincare} at side $(L+1)N'$ with
$c_0 \le \min(1,a)/\CP(L)$.  Then $c_0\le 0$.
\end{theorem}

\begin{proof}
If $c_0>0$, choose (Archimedes) $n$ with $n > \min(1,a)/c_0$.  By
Corollary~\ref{cor:linear}, $\CP(n)\ge n+1 > \min(1,a)/c_0$, whence
$\min(1,a)/\CP(n) < c_0$, contradicting domination at volume $n+1$.
\end{proof}

\begin{theorem}[the gate is false;
\lean{volumeUniformFlatPoincareGate\_false}]\label{thm:gatefalse}
For $d\ge3$, $N_c\ge2$, there is no single $\CP>0$ satisfying
\eqref{eq:poincare} at every volume $L+1$:
the proposition \lean{VolumeUniformFlatPoincareGate} --- with the
constant quantified \emph{before} the volume --- is provably false.
\end{theorem}

\begin{remark}[non-vacuity of the killed route;
\lean{perVolume\_flatPoincare\_family}]
The hypothesis family of Theorem~\ref{thm:noc} is inhabited: per-volume
constants exist at every volume.  The obstruction closes a route that is
genuinely open at each fixed volume.
\end{remark}

\begin{remark}[quantifier discipline]
The gate places its constant before the volume quantifier.  The reversed
order --- for every volume there exists a constant --- is exactly the
(true) per-volume statement, and conflating the two is the standard way
such obstructions get lost in prose.  The formalization makes the
distinction type-level.
\end{remark}

\section{The constant-sector quotient interface}\label{sec:quotient}

The wall witness lives in one sector.  The natural continuation
restricts the Poincar\'e question to its orthogonal complement --- the
physical fluctuation space.  The repository formalizes the
\emph{interface} (\lean{PhysicalPoincareSectorQuotient.lean}); we use
the word ``interface'' deliberately: no quotient type or packaged
orthogonal-complement submodule is constructed, and the predicate is a
restriction to fluctuation cochains.

\begin{definition}[fluctuation cochains; \lean{IsFluctuationCochain}]
$A$ is a fluctuation cochain if $\inn{A_v}{A}=0$ for every constant
generator $A_v$ --- equivalently, $A$ is orthogonal to the span of the
constant sector.  (The generator-wise form is chosen for formal
robustness.)
\end{definition}

\begin{proposition}[interface facts]
\raggedright
\emph{(i)} The constant sector is flat-harmonic
(\path{constantSectorLin_harmonic}) --- which is why only the block
term sees it, with the normalization of Lemma~\ref{lem:sector}.
\emph{(ii)} The full predicate \eqref{eq:poincare} implies its
restriction to fluctuation cochains at the same constant
(\path{quotientFlatPoincare_of_flatPoincare}); the restricted
statement is strictly weaker, so Theorem~\ref{thm:gatefalse} does
\emph{not} refute the quotient gate.
\emph{(iii)} Restricted constants exist at every fixed volume
(\path{exists_quotientFlatPoincare}).
\emph{(iv) (Non-transfer.)}  A nonzero constant cochain is not a
fluctuation cochain (\path{constant_not_fluctuation}): the wall
\emph{witness} is excluded from the fluctuation space by construction.
\end{proposition}

\begin{remark}[what the interface alone establishes]
\raggedright
At the interface stage, the volume-uniform quotient gate
(\path{VolumeUniformQuotientPoincareGate}) was neither proved nor
refuted: non-transfer removes the W-1 witness, not the possibility of a
second wall.  The next section supplies that wall with a different,
genuine fluctuation witness.  This chronology matters logically: the
W-1 obstruction is not being reused outside its domain.
\end{remark}

\section{The low-mode falsifier and the second wall}\label{sec:falsifier}

{\raggedright
A \emph{one-sided falsifier} for the quotient gate needs an explicit
fluctuation family on which the quadratic form of \eqref{eq:poincare}
can be evaluated.  The repository constructs the family
(\lean{PhysicalPoincareLowModeFalsifier.lean}), evaluates its Hodge term
exactly (\lean{PhysicalPoincareLowModeHodge.lean}), transports it to the
precise fine side consumed by the block map, and closes the Rayleigh
estimate and endpoint (\lean{PhysicalPoincareLowModeBlock.lean}).\par}

\begin{definition}[the half-period (two-interface) square-wave mode;
\lean{squareModeCochain}]
On even side lengths $N=2M$ ($M\ge1$), with oscillation direction $j$,
bond direction $i$, and internal vector $w$:
\[
A(x,k) \;=\;
\begin{cases}
\varepsilon(x_j)\, w, & k=i,\\[2pt]
0, & k\neq i,
\end{cases}
\qquad
\varepsilon(m)=\begin{cases}+1,& m<M,\\ -1,& m\ge M,\end{cases}
\]
(\lean{squareSign}).  The square profile is chosen over the cosine so
that every identity below is an exact finite $\pm1$ computation; odd
side lengths are out of scope by declared parameterization (a falsifier
needs only \emph{some} volume sequence).
\end{definition}

\begin{proposition}[exact identities]
\raggedright
\emph{(i)} $\sum_m \varepsilon(m) = 0$ (\lean{sum\_squareSign}), hence
$A$ is a fluctuation cochain
(\path{squareModeCochain_isFluctuation}) --- exact orthogonality, no
approximation.
\emph{(ii)} $\|A\|^2 = (2M)^d\,\|w\|^2$ exactly
(\path{norm_sq_squareModeCochain}).
\emph{(iii)} For $w\neq0$ the mode equals no constant cochain
(\path{squareModeCochain_ne_constant}), and the fluctuation space is
nontrivial for $N_c\ge2$ (\path{exists_nonzero_fluctuation}).
\end{proposition}

For arbitrary positive coarse side $N'$, the base witness with
half-period $MN'$ lives on the side
$(MN')+(MN')$, whereas the physical block predicate expects
$(M+M)N'$.  These sides are propositionally equal but their cochain types
carry size-indexed instances.  The formalization therefore uses canonical
coordinate, bond, site, and plaquette equivalences, lifted to linear
isometries, rather than an opaque dependent cast.  The resulting
\lean{blockScaleSquareModeCochain} preserves fluctuation orthogonality,
norm, shifts, flat curl, flat divergence, and Hodge energy.

\begin{theorem}[exact transported Hodge ratio]
\label{thm:hodge2}
\emph{Lean:}
\path{flatGaugeHodgeK0_inner_blockScaleSquareModeCochain_eq_div}.
Let $A$ be the transported mode on fine side $(2M)N'$, with $M,N'>0$.
For every adjoint model,
\begin{equation}\label{eq:hodge2}
 \inn{A}{K_0A}
 = \frac{8}{(2M)N'}\,\|A\|^2.
\end{equation}
The constant $8$ is the exact cost of the two jump interfaces.  For the
longitudinal mode it lies in the divergence term; for a transverse mode
it lies in the curl term.
\end{theorem}

\begin{theorem}[physical block contraction;
\lean{norm\_sq\_flatBlockConstraintQCLM\_le\_inv\_mul}]
\label{thm:blockcontract}
For every physical cochain $A$ and $d\ge3$,
\begin{equation}\label{eq:blockcontract}
 \|QA\|^2 \le \frac{1}{L}\,\|A\|^2.
\end{equation}
This is the current unscaled line-integral map measured in the current
unweighted coarse norm; no field rescaling or hidden volume weight is
inserted.
\end{theorem}

Taking $L=2M$ in Theorem~\ref{thm:blockcontract}, combining with
\eqref{eq:hodge2}, and using $N'\ge1$ yields the certified Rayleigh
numerator bound
\begin{equation}\label{eq:rayleigh2}
 \inn{A}{K_0A}+\|QA\|^2
 \le \frac{9}{2M}\,\|A\|^2
 \qquad
 (\lean{blockScaleSquareMode\_rayleigh\_numerator\_le}).
\end{equation}
The Hodge contribution is exact; the block contribution is the general
upper bound \eqref{eq:blockcontract}, and no pointwise equality for it is
claimed.

\begin{theorem}[linear lower bound in the fluctuation sector]
\label{thm:quotlower}
\emph{Lean:} \path{quotientPoincare_squareMode_linear_lower_bound}.
Let $M,N'>0$, $d\ge3$, $N_c\ge2$, and suppose the quotient Poincar\'e
predicate holds at block scale $2M$ with constant $\CP$.  Then
\[
 \frac{2M}{9}\le \CP.
\]
\end{theorem}

\begin{proof}
Choose a unit internal vector (available because $N_c\ge2$) and the
longitudinal transported mode.  It is a fluctuation cochain and has
strictly positive norm.  Apply the quotient predicate, use
\eqref{eq:rayleigh2}, and cancel $\|A\|^2>0$.
\end{proof}

\begin{theorem}[the quotient gate is false;
\lean{volumeUniformQuotientPoincareGate\_false}]
\label{thm:quotgatefalse}
For every positive $N'$, every $d\ge3$, every $N_c\ge2$, and every
\lean{SUNAdjointModel Nc},
\[
 \neg\,\lean{VolumeUniformQuotientPoincareGate}\ d\ N'\ N_c\ \rho.
\]
\end{theorem}

\begin{proof}
A putative gate supplies one constant $\CP$ before the block-scale
quantifier.  Theorem~\ref{thm:quotlower} forces $2M/9\le\CP$ for every
positive $M$.  Choosing $M$ by the Archimedean property gives a
contradiction.  The even subsequence suffices against the universal
block-scale quantifier.
\end{proof}

\begin{remark}[one-sided discipline and surviving alternatives]
The falsifier has fired in its conclusive direction: the diverging lower
bound refutes the gate, rather than merely eliminating a candidate.  Had
the quotient remained bounded on this mode, the gate would still not
have followed without an all-modes or spectral-minimality argument.  The
theorem is specific to the current unscaled $Q$ and current unweighted
coarse norm.  A rescaled or weighted block operator defines a different
gate and remains open, as does coercivity from the interacting Hessian.
\end{remark}

\section{Theorem--artifact map}\label{sec:map}

\begingroup
\footnotesize
\begin{longtable}{@{}>{\raggedright\arraybackslash}p{0.22\textwidth}>{\raggedright\arraybackslash}p{0.49\textwidth}>{\raggedright\arraybackslash}p{0.21\textwidth}@{}}
\toprule
Statement & Lean name & Module (\lean{YangMills/RG/}) \\
\midrule
\endfirsthead
\toprule
Statement & Lean name & Module (\lean{YangMills/RG/}) \\
\midrule
\endhead
CT endpoint, fixed volume & \leanprooflink{PhysicalShellCombesThomasEndpoint}{flatGaugeFixedCovariance_CT_fixedVolume} & \texttt{PhysicalShell}\newline\texttt{CombesThomasEndpoint} \\
CT positive-rate witness & \leanprooflink{PhysicalShellCombesThomasEndpoint}{exists_flatGaugeFixedCovariance_CT_fixedVolume} & \texttt{PhysicalShell}\newline\texttt{CombesThomasEndpoint} \\
$\dim_\R \su(n)=n^2-1$ & \leanprooflink{SUNAdjointDimension}{finrank_suLie} & \texttt{SUNAdjoint}\newline\texttt{Dimension} \\
Isometric transport & \leanprooflink{SUNAdjointModelInstance}{suLieCoordIso} & \texttt{SUNAdjoint}\newline\texttt{ModelInstance} \\
True adjoint model & \leanprooflink{SUNAdjointModelInstance}{matrixSUNAdjointModel} & \texttt{SUNAdjoint}\newline\texttt{ModelInstance} \\
Sector witness ($N_c\ge2$) & \leanprooflink{PhysicalPoincareWall}{exists_pos_constantSector} & \texttt{Physical}\newline\texttt{PoincareWall} \\
Wall, fixed volume & \leanprooflink{PhysicalPoincareWall}{flatPoincare_constantSector_lower_bound} & \texttt{Physical}\newline\texttt{PoincareWall} \\
Wall, linear form ($d\ge3$) & \leanprooflink{PhysicalPoincareWall}{flatPoincare_linear_lower_bound} & \texttt{Physical}\newline\texttt{PoincareWall} \\
Coercivity upper bound & \leanprooflink{PhysicalPoincareWall}{flatPoincare_coercivity_upper_bound} & \texttt{Physical}\newline\texttt{PoincareWall} \\
No volume-uniform $c_0$ & \leanprooflink{PhysicalPoincareWall}{no_volumeUniform_coercivity_via_flatPoincare} & \texttt{Physical}\newline\texttt{PoincareWall} \\
Gate false ($d\ge3$, $N_c\ge2$) & \leanprooflink{PhysicalPoincareWall}{volumeUniformFlatPoincareGate_false} & \texttt{Physical}\newline\texttt{PoincareWall} \\
Per-volume non-vacuity & \leanprooflink{PhysicalPoincareWall}{perVolume_flatPoincare_family} & \texttt{Physical}\newline\texttt{PoincareWall} \\
Fluctuation predicate & \leanprooflink{PhysicalPoincareSectorQuotient}{IsFluctuationCochain} & \texttt{PhysicalPoincare}\newline\texttt{SectorQuotient} \\
Full $\Rightarrow$ quotient & \leanprooflink{PhysicalPoincareSectorQuotient}{quotientFlatPoincare_of_flatPoincare} & \texttt{PhysicalPoincare}\newline\texttt{SectorQuotient} \\
Non-transfer of the witness & \leanprooflink{PhysicalPoincareSectorQuotient}{constant_not_fluctuation} & \texttt{PhysicalPoincare}\newline\texttt{SectorQuotient} \\
Quotient gate predicate & \leanprooflink{PhysicalPoincareSectorQuotient}{VolumeUniformQuotientPoincareGate} & \texttt{PhysicalPoincare}\newline\texttt{SectorQuotient} \\
Square profile, zero sum & \leanprooflink{PhysicalPoincareLowModeFalsifier}{sum_squareSign} & \texttt{PhysicalPoincare}\newline\texttt{LowModeFalsifier} \\
Mode is fluctuation & \leanprooflink{PhysicalPoincareLowModeFalsifier}{squareModeCochain_isFluctuation} & \texttt{PhysicalPoincare}\newline\texttt{LowModeFalsifier} \\
Mode norm, exact & \leanprooflink{PhysicalPoincareLowModeFalsifier}{norm_sq_squareModeCochain} & \texttt{PhysicalPoincare}\newline\texttt{LowModeFalsifier} \\
Exact transported Hodge ratio & \leanprooflink{PhysicalPoincareLowModeBlock}{flatGaugeHodgeK0_inner_blockScaleSquareModeCochain_eq_div} & \texttt{PhysicalPoincare}\newline\texttt{LowModeBlock} \\
Physical block contraction & \leanprooflink{PhysicalPoincareLowModeBlock}{norm_sq_flatBlockConstraintQCLM_le_inv_mul} & \texttt{PhysicalPoincare}\newline\texttt{LowModeBlock} \\
Rayleigh numerator bound & \leanprooflink{PhysicalPoincareLowModeBlock}{blockScaleSquareMode_rayleigh_numerator_le} & \texttt{PhysicalPoincare}\newline\texttt{LowModeBlock} \\
Quotient linear lower bound & \leanprooflink{PhysicalPoincareLowModeBlock}{quotientPoincare_squareMode_linear_lower_bound} & \texttt{PhysicalPoincare}\newline\texttt{LowModeBlock} \\
Quotient gate false & \leanprooflink{PhysicalPoincareLowModeBlock}{volumeUniformQuotientPoincareGate_false} & \texttt{PhysicalPoincare}\newline\texttt{LowModeBlock} \\
\bottomrule
\end{longtable}
\endgroup

Every listed name is exercised by the committed axiom-oracle script.  The
headline wall theorems, including all five new W-3c rows above, print
exactly \lean{[propext, Classical.choice, Quot.sound]} on the frozen tree.

\section{Reproducibility}\label{sec:repro}

\paragraph{Repository.}
\url{https://github.com/lluiseriksson/THE-ERIKSSON-PROGRAMME}
(branch \lean{main}).  The Lean tree described here is that of source
checkpoint
\href{https://github.com/lluiseriksson/THE-ERIKSSON-PROGRAMME/tree/a17d78165eaf7fc9ad505e8d3ed8e544848f8518}{\lean{a17d7816}},
sealed together with its transcript and live documentation by commit
\href{https://github.com/lluiseriksson/THE-ERIKSSON-PROGRAMME/commit/33072662b8d42c91185b852592bff51a6542b222}{\lean{33072662}};
this manuscript revision is
added in a strictly editorial later commit that changes no Lean file.

\paragraph{Toolchain.}  Lean \lean{leanprover/lean4:v4.29.0-rc6};
Mathlib pinned to commit
\lean{0764272048}\ldots{} (lakefile and manifest agree).  Build:
\lean{lake build YangMillsCore} --- expected
\lean{Build completed successfully (8412 jobs)}.

\paragraph{Axiom oracle.}  \lean{lake env lean oracle\_check.lean}
executes $2250$ \lean{\#print axioms} invocations ($2247$ distinct output
names); the committed
\href{https://github.com/lluiseriksson/THE-ERIKSSON-PROGRAMME/blob/3923b93272ac9bf3b13de5e1539b522ddbbcbc95/docs/oracle-transcripts/ORACLE-20260714-a17d7816.txt}{transcript
\lean{ORACLE-20260714-a17d7816.txt}} records the
full raw output with provenance header: $2228$ results on a nonempty
subset of \lean{[propext, Classical.choice, Quot.sound]} (finely
$2145+65+18$) and $22$ axiom-free results; zero \lean{sorryAx}, zero
nonstandard axiom sets.  SHA-256 hashes (git blob form, LF):
script \lean{CBEFB14D}\ldots, raw output \lean{4176F9E9}\ldots.

\paragraph{Verification ledger.}  The append-only
\href{https://github.com/lluiseriksson/THE-ERIKSSON-PROGRAMME/blob/3923b93272ac9bf3b13de5e1539b522ddbbcbc95/docs/VERIFICATION-LEDGER.md}{verification ledger}
documents every checkpoint of the
results here (Addenda 467--507), including the measurement-discipline
rules (measure the pushed tree; blob-hash artifacts; positive-assertion
counter sweeps) and the corrections history.

\section*{Continuation from this submission}

The Hodge evaluation, block estimate, Rayleigh bound, and endpoint
decision are now Theorems~\ref{thm:hodge2}--\ref{thm:quotgatefalse}, not
future work.  The surviving registered routes are: a rescaled or weighted
block operator with its normalization re-audited from first principles,
and coercivity extracted from the interacting Wilson Hessian.  No claim
that either route succeeds is made.  Any continuation will retain the
same discipline --- exact identities, declared parameterizations, and
one-sided claims only where only one side is proved.

\begin{thebibliography}{99}

\bibitem{Balaban84a}
T.~Ba{\l}aban,
\emph{Propagators and renormalization transformations for lattice gauge
theories.~I},
Comm.\ Math.\ Phys.\ \textbf{95} (1984) 17--40.

\bibitem{Balaban84b}
T.~Ba{\l}aban,
\emph{Propagators and renormalization transformations for lattice gauge
theories.~II},
Comm.\ Math.\ Phys.\ \textbf{96} (1984) 223--250.

\bibitem{CombesThomas}
J.~M.~Combes and L.~Thomas,
\emph{Asymptotic behaviour of eigenfunctions for multiparticle
Schr\"odinger operators},
Comm.\ Math.\ Phys.\ \textbf{34} (1973) 251--270.

\bibitem{KoteckyPreiss}
R.~Koteck\'y and D.~Preiss,
\emph{Cluster expansion for abstract polymer models},
Comm.\ Math.\ Phys.\ \textbf{103} (1986) 491--498.

\bibitem{JaffeWitten}
A.~Jaffe and E.~Witten,
\emph{Quantum Yang--Mills theory},
Clay Mathematics Institute Millennium Prize problem description (2000).

\bibitem{Lean4}
L.~de~Moura and S.~Ullrich,
\emph{The Lean 4 theorem prover and programming language},
CADE-28, LNCS 12699 (2021) 625--635.

\bibitem{mathlib}
The mathlib community,
\emph{The Lean mathematical library},
CPP 2020, 367--381.

\bibitem{ErikssonVU}
L.~Eriksson,
\emph{A machine-checked volume-uniform Wilson-loop area law via a
formalized cluster expansion},
\href{https://ai.vixra.org/abs/2607.0005}{ai.viXra.org:2607.0005} (2026).

\bibitem{ErikssonHoles}
L.~Eriksson,
\emph{Machine-checked rooted-tree majorants for polymer expansions with
holes},
\href{https://ai.vixra.org/abs/2607.0025}{ai.viXra.org:2607.0025} (2026).

\bibitem{ErikssonC4}
L.~Eriksson,
\emph{A machine-checked proof of an Amos-type bound for modified Bessel
ratios at real order},
\href{https://ai.vixra.org/abs/2607.0030}{ai.viXra.org:2607.0030} (2026).

\bibitem{ErikssonC3}
L.~Eriksson,
\emph{A machine-checked proof of the Amos bound for modified Bessel
function ratios},
\href{https://ai.vixra.org/abs/2607.0032}{ai.viXra.org:2607.0032} (2026).

\bibitem{ErikssonC2}
L.~Eriksson,
\emph{One Amos bound, three consumer sites: machine-checked
Bessel-ratio calculus for lattice gauge expansions},
\href{https://ai.vixra.org/abs/2607.0033}{ai.viXra.org:2607.0033} (2026).

\bibitem{Eriksson2D}
L.~Eriksson,
\emph{Exact two-dimensional $\SU(2)$ Yang--Mills in Lean: Weyl
integration, heat-kernel convolution, Migdal invariance, and the exact
simple-loop area law},
\href{https://ai.vixra.org/abs/2607.0035}{ai.viXra.org:2607.0035} (2026).

\bibitem{ErikssonCatalan}
L.~Eriksson,
\emph{A machine-verified bijective proof of the rooted child-factorial
Catalan identity over spanning trees of the complete graph},
\href{https://ai.vixra.org/abs/2607.0001}{ai.viXra.org:2607.0001} (2026).

\bibitem{ErikssonMap}
L.~Eriksson,
\emph{The master map --- audit experiments report: mechanical audit
experiments and reproducibility appendix for the 2602-series programme
on 4D $\SU(N)$ Yang--Mills},
\href{https://ai.vixra.org/abs/2602.0117}{ai.viXra.org:2602.0117} (2026).

\end{thebibliography}

\end{document}
