\documentclass[11pt,a4paper]{article}
\usepackage[margin=1.05in]{geometry}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{booktabs}
\usepackage{microtype}
\usepackage{enumitem}
\usepackage[colorlinks=true,linkcolor=blue,urlcolor=blue,citecolor=blue]{hyperref}
\hypersetup{
  pdftitle={Machine-Checked CMP116 Fluctuation Reduction: Physical Constraint Coordinates and the Interacting-Hessian Frontier},
  pdfauthor={Lluis Eriksson},
  pdfsubject={Lean 4 formalization of the CMP116 finite Gaussian reduction}
}

\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{remark}[theorem]{Remark}

\emergencystretch=3em
\tolerance=1800

\newcommand{\lean}[1]{\texttt{#1}}
\newcommand{\SU}{\mathrm{SU}}
\newcommand{\Tr}{\operatorname{Tr}}
\newcommand{\opnorm}[1]{\lVert #1\rVert_{2\to2}}
\newcommand{\gauss}{\gamma}
\newcommand{\cov}{\mathcal C}
\newcommand{\proj}{P_{Z_0}}

\title{Machine-Checked CMP116 Fluctuation Reduction:\\
Physical Constraint Coordinates and the Interacting-Hessian Frontier}
\author{Lluis Eriksson\thanks{This manuscript and parts of the formal
development were produced with AI assistance under the repository's audit
protocol. Every result labelled \emph{machine-checked} is a Lean~4 theorem
against a pinned Mathlib. Repository:
\url{https://github.com/lluiseriksson/THE-ERIKSSON-PROGRAMME}.}}
\date{July 2026}

\begin{document}
\maketitle

\begin{abstract}
We give a machine-checked reduction of the finite-dimensional fluctuation
integral in Ba\l aban's CMP116 large-field analysis.  Starting from the
physical block constraint $Q$, the formal development constructs a sparse
right inverse $E$ and the constraint-elimination operator
$C=I-EQ$.  It proves $QE=I$, $QC=0$, $C^2=C$, the exact sparse norm
$\lVert EB\rVert=M^{d-1}\lVert B\rVert$, and the volume-independent bound
$\lVert C\rVert\le 1+M^{d-1}$ for $d\ge3$.  An exact physical/CMP116
isometry transports $C$ to finite Gaussian coordinates without norm loss.
The same development constructs the physical localization projector
$P_{Z_0}$, evaluates the complex quadratic Gaussian, localizes its determinant
to rank $|I(Z_0)|$, performs the outer Gaussian integration, and absorbs both
costs into an explicit $\exp(c|Z_0|)$ factor.

Two corrections exposed by formalization are central.  First, the useful
domination occurs after Gaussian integration rather than through an
unavailable pointwise supremum in the fluctuation field.  Second, the
localized quadratic matrix is $A=-\alpha_5P_{Z_0}$.  In the exactly identified
trivial-background sector the terminal Lean theorem inserts the concrete
$C$, the flat Hessian, the complement localization, and the covariance root
directly into the printed source
$\Gamma_k=C^T\Delta_k(CP_{Z_0^c})(C^{(k)})^{1/2}$ and returns an explicit
Cauchy bound with no ambient-volume factor.  CMP116, however, requires the
base Hessian at a generally nontrivial small background $\bar U$.  We do not
construct $D^2S_{\rm Wilson}(\bar U)$ or the random-walk estimate (2.16), and
therefore do not prove the physical domination, (2.26), \lean{hraw},
\lean{hRpoly}, a continuum limit, or a mass gap.  The contribution is an
auditable reduction that closes the constraint and Gaussian layers and
identifies the first genuinely missing interacting construction.
\end{abstract}

\section{Introduction}

Rigorous renormalization-group analyses of lattice gauge theory separate
small- and large-field regions, integrate finite-dimensional fluctuation
fields, and reorganize the result into polymer activities.  Ba\l aban's
series~\cite{Bal87,Bal88a,Bal88b,Bal89} and Dimock's exposition
\cite{Dim13a,Dim13b,Dim14} provide the mathematical source for the present
formalization.  The specific target is the fluctuation-integral step around
equations (2.14), (2.20)--(2.26) of the CMP116 large-field analysis.

The purpose of the development is not to replace an analytic estimate by an
interface bearing the same estimate as a hidden hypothesis.  Instead, it
asks which parts of the source argument are finite, algebraic, geometric, or
measure-theoretic and can therefore be closed independently, and which part
is the genuine remaining Yang--Mills estimate.

The machine-checked chain established here is
\begin{equation}\label{eq:chain}
\begin{aligned}
 (D,P)&\longmapsto Z_0\longmapsto P_{Z_0}
 \\[-1mm]
 Q&\longmapsto E\longmapsto C=I-EQ
 \longmapsto C_{\mathrm{matrix}}\\
 (C_{\mathrm{matrix}},\Delta_{\mathrm{flat}},R_{\mathrm{matrix}})
 &\longmapsto \Gamma_{\mathrm{flat}}
 \longmapsto \text{localized Gaussian cost}
 \longmapsto \text{Cauchy term bound}.
\end{aligned}
\end{equation}
Here $D$ denotes the family of large-field components, $P$ the selected
large-field bonds, $Z_0$ their least admissible localization region,
$R_{\mathrm{matrix}}$ is the finite matrix of the physical covariance root,
while $C_{\mathrm{matrix}}$ is the transported constraint elimination.  The
terminal theorem in the flat sector leaves the physical domination as its
substantive input; all source-norm, determinant, cardinality, and outer
Gaussian costs are produced internally from explicit operator bounds.

This separation is useful even before those estimates are proved.  It prevents
three common sources of ambiguity: choosing the localization projector by
hand, placing an impossible supremum before a Gaussian integration, and using
the wrong sign in the shifted quadratic form.

\section{Finite physical localization}

Let $Z_0$ be a finite family of coarse blocks.  A physical site belongs to the
interior when the site and all its forward and backward nearest neighbours
belong to the union of blocks.  A positively oriented physical bond is
interior when both endpoints are interior.  This two-endpoint convention is
important: a one-endpoint convention does not represent the source boundary
condition faithfully.

The development constructs a least region
\[
  Z_0=\operatorname{LocalizationCore}(D,P)
\]
that contains the geometric seed induced by $D$ and makes every $e\in P$
interior.  It proves the universal property
\begin{equation}\label{eq:minimal}
 \operatorname{LocalizationCore}(D,P)\subseteq W
 \quad\Longleftrightarrow\quad
 W\text{ contains the seed and every }e\in P\text{ is interior to }W.
\end{equation}
The principal Lean endpoints are
\lean{cmp116LocalizationCore\_subset\_iff\_admissible} and
\lean{cmp116LocalizationCore\_admissible}.

The physical-to-scalar dictionary contains an equivalence
\[
 \operatorname{coordEquiv}: (q,a)\simeq(e,b)
\]
between CMP116 scalar coordinates and physical bond/Lie coordinates.  It
defines the canonical finite carrier
\begin{equation}\label{eq:carrier}
 I(Z_0)=\{qa:\operatorname{bond}(\operatorname{coordEquiv}(qa))
                  \text{ is interior to }Z_0\}.
\end{equation}
The projector $P_{Z_0}$ is the diagonal matrix with entries $1$ on $I(Z_0)$
and $0$ elsewhere.  The following theorem is machine-checked.

\begin{theorem}[Physical localization projector]\label{thm:projector}
For every scalar field $x$ and coordinate $qa$,
\[
 (P_{Z_0}x)_{qa}=\begin{cases}
 x_{qa},&qa\in I(Z_0),\\
 0,&qa\notin I(Z_0),
 \end{cases}
 \qquad
 x^TP_{Z_0}x=\sum_{qa\in I(Z_0)}x_{qa}^2.
\]
If $Z_0$ is admissible for $(D,P)$, then for every $e\in P$ and every Lie
coordinate $a$, $\operatorname{coordEquiv}^{-1}(e,a)\in I(Z_0)$.  The same
holds for the least region $\operatorname{LocalizationCore}(D,P)$.
\end{theorem}

This is formalized in
\lean{BalabanCMP116Eq223PhysicalLocalizationProjector.lean}.  In particular,
the shifted Hessian below no longer accepts an arbitrary user-supplied set of
coordinates.

\section{Physical constraint elimination}

Let $Q$ be the flat block constraint from fine one-cochains to coarse
one-cochains.  For every coarse bond the formalization chooses one
distinguished fine pivot bond and defines a sparse insertion $E$ supported on
those pivots.  The normalization is fixed so that
\begin{equation}\label{eq:rightinverse}
 QE=I.
\end{equation}
The physical elimination operator is then
\begin{equation}\label{eq:Cdef}
 C=I-EQ.
\end{equation}
This is the padded square realization of the rectangular free-coordinate map:
on the full fine space it is the projection onto $\ker Q$, while on inputs
vanishing at the pivot bonds it parametrizes the constrained field by the
remaining bonds.  The distinction matters; no Gaussian Jacobian statement is
inferred merely from the padded realization.

\begin{theorem}[Constraint elimination and uniform norm]
\label{thm:constraint}
The operators $E,Q,C$ satisfy
\[
 QE=I,\qquad QC=0,\qquad
 Cx=x\ (Qx=0),\qquad C^2=C.
\]
Every nonpivot coordinate is preserved, and each pivot coordinate is changed
by exactly the coarse constraint value.  Moreover
\begin{equation}\label{eq:Cnorm}
 \lVert EB\rVert=M^{d-1}\lVert B\rVert,
 \qquad
 \lVert C\rVert\le 1+M^{d-1}\quad(d\ge3),
\end{equation}
independently of the periodic volume $N'$.
\end{theorem}

The first equality in \eqref{eq:Cnorm} follows from the disjoint support of
the pivot insertion.  The second combines it with the certified contraction
of $Q$ and the triangle inequality in $C=I-EQ$.  The exact isometric
physical/CMP116 dictionary defines $C_{\rm matrix}$ and proves
\begin{equation}\label{eq:Ctransport}
 \lVert C_{\rm matrix}\rVert=\lVert C\rVert,
 \qquad
 \operatorname{coordinates}(C_{\rm matrix}x)
   =C(\operatorname{coordinates}(x)).
\end{equation}
Thus the strongest flat-sector endpoint receives neither an arbitrary matrix
$C$, nor a constant $C_{\rm norm}$, nor a proof bounding it.

\section{Exact Gaussian calculus and physical covariance}

Let $\gauss$ denote the standard Gaussian probability measure on a finite
real coordinate space.  For a real symmetric matrix $A$ with $I+A>0$ and a
complex source $c$, the formal development proves
\begin{equation}\label{eq:gaussian}
 \int \exp\!\left(-\frac12x^TAx+c^Tx\right)d\gauss(x)
 =\frac{1}{\sqrt{\det(I+A)}}
   \exp\!\left(\frac12c^T(I+A)^{-1}c\right).
\end{equation}
The source pairing is bilinear, not Hermitian: there is no complex conjugation
on $c$.  The proof proceeds through the one-dimensional integral, finite
products, orthogonal invariance, spectral diagonalization, determinant
reconstruction, and the inverse quadratic form.  Coupled two-stage integrals
and their Schur complement are also formalized.

For a real covariance-root matrix $R$, the pushforward measure
$R_*\gauss$ is a positive probability measure.  Substituting $x=Ru$ in
\eqref{eq:gaussian} gives the majorant
\begin{equation}\label{eq:majorant}
 M_{2.24}(R,A,c)=
 \det(I+R^TAR)^{-1/2}
 \exp\!\left(\frac12\operatorname{Re}
 \bigl[c^T(I+R^TAR)^{-1}c\bigr]\right).
\end{equation}
The equality between the norm of the complex Gaussian integral and
$M_{2.24}$ is machine-checked.  So is the domination principle
\begin{equation}\label{eq:domination}
 |f(u)|\le e^{-u^TAu/2+r^Tu}\ \text{a.e.}
 \quad\Longrightarrow\quad
 \left|\int f\,d(R_*\gauss)\right|\le M_{2.24}(R,A,r).
\end{equation}

The matrix part of this majorant is also quantitatively closed.  Put
$q=\alpha_5\opnorm R^2<1$, let $S=I(Z_0)$, and take
$A=-\alpha_5P_S$.  A rectangular Sylvester identity moves the determinant to
the localized coordinate subtype and yields
\begin{equation}\label{eq:rankdet}
 (1-q)^{|S|}
 \le \det(I-\alpha_5R^TP_SR).
\end{equation}
The exponent is the physical localized rank, not the dimension of the ambient
Gaussian space.  The inverse source form is bounded spectrally, while the
remaining outer Gaussian moment is evaluated exactly:
\begin{equation}\label{eq:outermoment}
 \int \exp\!\left(\beta\sum_{i\in S}x_i^2\right)d\gauss(x)
 =\bigl(1-2\beta\bigr)^{-|S|/2},\qquad 2\beta<1.
\end{equation}
Finally, the explicit geometric estimate
\begin{equation}\label{eq:Scard}
 |S|\le |Z_0|M^d d(N_c^2-1)
\end{equation}
absorbs both \eqref{eq:rankdet} and \eqref{eq:outermoment} into
\begin{equation}\label{eq:fullmajorant}
 \text{localized determinant cost}\times\text{outer moment}
 \le \exp\!\left[(c_{\det}+c_G)|Z_0|\right],
\end{equation}
where both rates are explicit logarithmic expressions in the corresponding
small parameters.  There is no hidden factor depending on the total volume.

The physical covariance certificate supplies an operator $C$, a root
$C^{1/2}$, self-adjointness, positivity, the exact identity
$(C^{1/2})^2=C$, and localization/norm bounds.  The scalar dictionary is
proved to be an exact $L^2$ isometry.  Therefore its matrix realization
$R_{\mathrm{matrix}}$ satisfies
\begin{equation}\label{eq:norms}
 \opnorm{R_{\mathrm{matrix}}}=\opnorm{C^{1/2}},
 \qquad
 \opnorm{R_{\mathrm{matrix}}}^{,2}=\opnorm{C}.
\end{equation}
There is no dimension-dependent condition number in this transport.

\section{Two corrections forced by formalization}

\subsection{Integrate before taking the norm}

A tempting interface asks for a pointwise bound
\[
 |\operatorname{innerIntegrand}(X,B)|\le M
 \quad\text{uniformly in }B.
\]
For the source bilinear factor $\exp(-\langle B,\Gamma X\rangle)$ this is
generally false: along a direction of $B$ with the appropriate sign, the real
part of the exponent grows.  Such a hypothesis would make a downstream
theorem formally true but physically unusable.

The repaired interface integrates against the conditioned Gaussian first:
\begin{equation}\label{eq:integrated}
 \left|\int \operatorname{innerIntegrand}(X,B),d\mu_{\sigma,\tau}(B)\right|
 \le M_{\mathrm{inner}}.
\end{equation}
Equation~\eqref{eq:domination}, followed by the exact evaluation
\eqref{eq:majorant}, produces this bound while retaining the Gaussian damping
and completing the square.  The outer integration is then controlled by its
probability measure and the outer weight.

\subsection{The localized quadratic matrix has negative sign}

The localized source factor contains
\[
 \exp\!\left(\frac{\alpha_5}{2}\lVert P_{Z_0}B\rVert^2\right).
\]
In the convention of \eqref{eq:gaussian}, this corresponds to
\begin{equation}\label{eq:sign}
 A=-\alpha_5P_{Z_0},
\end{equation}
not to a positive-semidefinite $A$.  Hence the actual positivity obligation is
\begin{equation}\label{eq:shifted}
 I-\alpha_5R_{\mathrm{matrix}}^TP_{Z_0}R_{\mathrm{matrix}}>0.
\end{equation}
By \eqref{eq:norms}, the source condition
\begin{equation}\label{eq:small}
 0\le\alpha_5,
 \qquad \alpha_5\,\mathrm{covNormBound}<\frac12,
 \qquad \opnorm C\le\mathrm{covNormBound},
\end{equation}
implies \eqref{eq:shifted}.  The corresponding Lean endpoint is
\begin{center}\small
\texttt{posDef\_physicalRootMatrix\_of}\\
\texttt{\_alpha5\_covariance\_half\_small}\\
\texttt{\_physicalZ0}.
\end{center}
Its canonical-core specialization takes only $(D,P)$ as localization data.

\section{The main reduction theorem}

Let $G$ denote the finite CMP116 data for the contour variables, weights,
cutoffs, field evaluation, and interaction exponent.  In the exactly
identified trivial-background sector define
\begin{align}
 \Delta_{\rm flat}&=K_0+aQ^*Q,\label{eq:flatdelta}\\
 \Gamma_{\rm flat}
 &=C_{\rm matrix}^T\Delta_{\rm flat}
   (C_{\rm matrix}P_{Z_0^c})R_{\rm matrix}.\label{eq:gammaflat}
\end{align}
The finite-range kernel and bond-ball count give the volume-independent bound
\begin{equation}\label{eq:deltabound}
 \opnorm{\Delta_{\rm flat}}
 \le \bigl((4d)^2+4+|a|M^2\bigr)
       \bigl(2(3M+1)\bigr)^d d.
\end{equation}
Since coordinate transport is isometric and
$\opnorm{CP_{Z_0^c}}\le\opnorm C$, equations
\eqref{eq:Cnorm} and \eqref{eq:deltabound} produce an explicit squared source
rate for \eqref{eq:gammaflat}.  The only substantive functional premise left
in the strongest endpoint is the physical Gaussian domination
\begin{equation}\label{eq:hdom}
 |\operatorname{innerIntegrand}_{G}(b)|
 \le
 \exp\!\left(
  \frac{\alpha_5}{2}b^TP_{Z_0}b
  +\langle \Gamma_{\rm flat}P_{Z_0}x,b\rangle
 \right)\qquad\text{(Dom)}
\end{equation}
for $(R_{\rm matrix})_*\gauss$-almost every $b$ on the Cauchy contours.

\begin{theorem}[Physical-constraint Gaussian reduction, flat sector]
\label{thm:main}
Assume $d\ge3$, a physical localized covariance-root certificate, the
smallness conditions needed for the two Gaussian moments, positive Cauchy
radii, a nonnegative uniform outer-weight bound $M_{\rm out}$, and
\eqref{eq:hdom}.  Then the literal finite equation-(2.14) term satisfies
\begin{equation}\label{eq:terminal}
 \lVert\operatorname{term}_{G}(Y_0,P;\psi,\phi)\rVert
 \le
 \operatorname{CauchyRate}_{\Delta}\!\left(
   \operatorname{CauchyRate}_{Y}\!\left(
    M_{\rm out}\exp(c_{\rm tot}|Z_0|)\right)\right),
\end{equation}
where $c_{\rm tot}$ is the explicit total Gaussian cardinality rate formed
from $M,d,N_c,\alpha_5$, the covariance-root norm, the bound
$1+M^{d-1}$ for $C$, and the right-hand side of
\eqref{eq:deltabound}.  No arbitrary $C$, $C$-norm certificate, Hessian matrix,
Hessian-norm certificate, ambient determinant dimension, uniform-in-$x$
source bound, or free Gaussian-majorant premise occurs in the endpoint.
\end{theorem}

The Lean endpoint is
\begin{center}\small
\nolinkurl{norm_term_le_cauchyRate_of_physicalConstraintGamma}\
\nolinkurl{\_flatDelta_expCard}
\end{center}
in \lean{BalabanCMP116Eq214PhysicalConstraintC.lean}.

\begin{remark}[What the theorem does not say]
Theorem~\ref{thm:main} is a reduction theorem.  It does not prove
the concrete interacting-background kernel hypotheses used to obtain
\eqref{eq:hdom}, the random-walk estimate (2.16), the subsequent termwise
factorization of equation (2.26), or \lean{hRpoly}.
Consequently it implies neither a lattice mass gap nor a continuum
Yang--Mills construction.
\end{remark}

\section{Formal architecture and audit}

Table~\ref{tab:map} maps the mathematical chain to the principal modules.
All entries abbreviate filenames with prefix \lean{BalabanCMP116} and suffix
\lean{.lean}.

\begin{table}[ht]
\centering
\small
\begin{tabular}{p{0.30\linewidth}p{0.61\linewidth}}
\toprule
Layer & Lean module \\
\midrule
Physical incidence and least $Z_0$ &
\lean{Eq214Interior},
\lean{Eq214LocalizationCore} \\
Cutoff and Cauchy extraction &
\lean{Eq23Cutoff},
\lean{Eq214Cauchy} \\
Exact Gaussian identities &
\lean{QuadraticGaussianMatrix},
\lean{CorrelatedGaussian} \\
Domination and (2.24) &
\lean{Eq223GaussianDomination},
\lean{Eq224GaussianBound},
\lean{Eq224DeterminantBound},
\lean{Eq224SourceBound} \\
Physical root and norm transport &
\lean{Eq223PhysicalRootNormBridge} \\
Physical $P_{Z_0}$ &
\lean{Eq223PhysicalLocalizationProjector} \\
Constraint right inverse and $C=I-EQ$ &
\lean{CMP96ConstraintRightInverse},
\lean{CMP96ConstraintElimination} \\
Uniform norm and CMP116 transport of $C$ &
\lean{CMP96ConstraintNorm},
\lean{Eq214PhysicalConstraintC} \\
Localized determinant, outer moment &
\lean{Eq224LocalizedDeterminant},
\lean{Eq225OuterGaussianMoment},
\lean{Eq226GaussianCardinality} \\
Physical source and flat Hessian &
\lean{Eq214GammaComplement},
\lean{Eq214FlatDeltaBound} \\
Terminal reduction &
\lean{Eq214PhysicalConstraintC} \\
\bottomrule
\end{tabular}
\caption{Principal theorem-to-artifact map.}
\label{tab:map}
\end{table}

The repository pins Lean and Mathlib.  The root build after the physical
constraint checkpoint completes 8,467 jobs.  The 32 new mathematical
endpoints across the construction, norm, and terminal-consumption checkpoints
were audited separately; each prints only \lean{propext},
\lean{Classical.choice}, and \lean{Quot.sound}.  The added modules contain no
\lean{sorry}, \lean{admit}, local \lean{axiom}, \lean{unsafe}, or
\lean{opaque} declaration.

The reproducible checkpoint is maintained on the repository branch
\lean{codex/hrpoly-b3-card-tilt}, pull request
\href{https://github.com/lluiseriksson/THE-ERIKSSON-PROGRAMME/pull/28}{\#28}.
The immutable release tag is
\lean{hrpoly-cmp116-main-reduction-v0.3-2026-07-16}.  The exact commands are
\begin{verbatim}
lake exe cache get
lake build YangMillsCore
lake env lean oracle_check.lean
\end{verbatim}

\section{Remaining analytic frontier}

The formalization localizes the remaining difficulty rather than reducing its
mathematical importance.  CMP116 does not identify the contour Hessian with
the flat operator \eqref{eq:flatdelta}.  Its base point is
$(\sigma(Z),U,J)=(0,\bar U,0)$, where $\bar U$ is a generally nontrivial small
background.  The differences in the quadratic forms, covariance, and mixed
source produce $R_1,R_2,R_3$.  The required estimate has the schematic form
\begin{equation}\label{eq:216}
 |R_i(b,b')|
 \le
 \left[c_0e^{-\delta_0M/3}+c_1(\alpha_0+\alpha_1)\right]
 e^{-\delta_0 d(b,b')/2}.
\end{equation}
The repository contains consumers of such a bound, but no load-bearing
construction of the second variation of the nonabelian Wilson action at
$\bar U$ and no random-walk expansion proving \eqref{eq:216}.

Thus the remaining route is
\[
\begin{aligned}
 S_{\rm Wilson}
 &\longrightarrow D^2S_{\rm Wilson}(\bar U)
 \longrightarrow \text{random-walk expansion}
 \longrightarrow (2.16)\\
 &\longrightarrow h_{\rm dom}
 \longrightarrow (2.26)
 \longrightarrow h_{\mathrm{raw}}
 \longrightarrow \mathrm{KP}.
\end{aligned}
\]
The determinant, inverse-source, cardinality, constraint-coordinate, and flat
source-norm analyses are already internal.  Introducing (2.16) as a renamed
hypothesis would not close this frontier; the next valuable endpoint must
derive it from the physical Wilson Hessian.

\section{Conclusion}

The formal development closes the finite geometry, physical constraint
elimination, coordinate transport, Gaussian evaluation, determinant
localization, outer integration, sign, and positivity steps between CMP116
localization data and an explicit polymer-volume Cauchy bound.  It also
records why three superficially convenient formulations are wrong: bounding
the bilinear inner factor pointwise before Gaussian integration, assigning a
positive sign to the localized quadratic matrix, and replacing the small
interacting background by the trivial background.  The terminal theorem is
both useful and deliberately incomplete: it exposes, without wrappers, the
physical Wilson-Hessian and random-walk estimates that still carry the
analytic content of (2.16), equation (2.26), and \lean{hRpoly}.

\begin{thebibliography}{99}
\small

\bibitem{Bal87} T.~Ba{\l}aban,
\emph{Renormalization group approach to lattice gauge field theories.
I. Generation of effective actions in a small field approximation and a
coupling constant renormalization in four dimensions},
Comm. Math. Phys. \textbf{109} (1987), 249--301.
doi:10.1007/BF01215223.

\bibitem{Bal88a} T.~Ba{\l}aban,
\emph{Renormalization group approach to lattice gauge field theories.
II. Cluster expansions}, Comm. Math. Phys. \textbf{116} (1988), 1--22.
doi:10.1007/BF01239022.

\bibitem{Bal88b} T.~Ba{\l}aban,
\emph{Convergent renormalization expansions for lattice gauge theories},
Comm. Math. Phys. \textbf{119} (1988), 243--285.
doi:10.1007/BF01217741.

\bibitem{Bal89} T.~Ba{\l}aban,
\emph{Large field renormalization. II. Localization, exponentiation, and
bounds for the R operation}, Comm. Math. Phys. \textbf{122} (1989), 355--392.
doi:10.1007/BF01238433.

\bibitem{Dim13a} J.~Dimock,
\emph{The renormalization group according to Balaban. I. Small fields},
Rev. Math. Phys. \textbf{25} (2013), 1330010.
doi:10.1142/S0129055X13300100; arXiv:1108.1335.

\bibitem{Dim13b} J.~Dimock,
\emph{The renormalization group according to Balaban. II. Large fields},
J. Math. Phys. \textbf{54} (2013), 092301.
doi:10.1063/1.4821275; arXiv:1212.5562.

\bibitem{Dim14} J.~Dimock,
\emph{The renormalization group according to Balaban. III. Convergence},
Ann. Henri Poincar\'e \textbf{15} (2014), 2133--2175.
doi:10.1007/s00023-013-0303-3.

\bibitem{mathlib20} The mathlib Community,
\emph{The Lean mathematical library}, Proc. CPP 2020, ACM, 2020,
367--381. doi:10.1145/3372885.3373824.

\end{thebibliography}

\end{document}
