\documentclass[11pt]{article}
\usepackage{cmap}
\usepackage[margin=1.1in]{geometry}
\usepackage{amsmath,amssymb,amsthm}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\input{glyphtounicode}
\pdfgentounicode=1
\usepackage{xcolor}
\usepackage{hyperref}
\definecolor{linkblue}{rgb}{0.1,0.2,0.55}
\hypersetup{colorlinks=true,linkcolor=linkblue,citecolor=linkblue,urlcolor=linkblue}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\theoremstyle{remark}
\newtheorem{remark}[theorem]{Remark}
\newtheorem{problem}[theorem]{Open problem}

\title{A Weighted Tur\'an-Type Monotonicity Lemma for Modified
Bessel Functions via the Calibrated Amos Bound, with an Application
to the Ordering of a Surface Expansion in Two-Dimensional Lattice
Gauge Theory}
\author{Lluis Eriksson}
\date{July 2026}

\begin{document}
\maketitle

\begin{abstract}
For $x > 0$ and real $m \ge 1$ define, with $I_\mu$ the modified
Bessel function of the first kind,
\[
\varphi_m(x) \;=\;
\frac{(m-1)\,I_{m-1}(x)^2 + (m+1)\,I_{m+1}(x)^2}{m\,I_m(x)^2}.
\]
We prove that $\varphi_m(x) < \varphi_{m+1}(x)$ for every $x>0$ and
every real $m \ge 1$ --- a weighted Tur\'an-type monotonicity
statement we have not found in the literature, although every
ingredient of the proof is classical. The proof is fully elementary
and uses exactly the arsenal of the companion note on unit-step
order-monotonicity: the three-term recurrence and the Amos-type
bound $\rho_m < x/(a + \sqrt{a^2+x^2})$, $a = m + \tfrac12$, for
$\rho_m = I_{m+1}/I_m$. The engine is a new exact factorization:
with $u = \rho_m(x)$, $c = 1/x$, $S = u + 1/u$, $P = 1/u - u$,
\[
\frac{m(m+1)\bigl(\varphi_{m+1}-\varphi_m\bigr)}{2m(m+1)}
\;=\; (S - 3c)\,\bigl(P - (2m+1)c\bigr) \;+\; (2m+1)c^2 ,
\]
an algebraic identity once $\rho_{m-1}$ and $\rho_{m+1}$ are
eliminated by the recurrence. The second factor is positive by the
calibrated Amos bound --- it is precisely the unit-step inequality of
the companion note --- and the first because the same bound forces
$u < x/(2m+1) \le x/3$. As an application we obtain the strict
ordering $c_{mn} < 0$ ($m<n$) of the $2\times2$ determinants of the
coefficient arrays of a $\pi$-local surface expansion arising in an
audit-first programme on two-dimensional $SU(2)$ lattice gauge
theory; this is the structural half of a positivity problem whose
remaining half --- an alternating-parity dominance with
near-perfect mirror cancellation --- we state precisely and leave
open. The full chain of lemmas is machine-checked in Lean~4/Mathlib
(no \texttt{sorry}; standard axiom oracle), with the analytic inputs
carried as named hypotheses, and independently audited numerically
at high precision.
\end{abstract}

\section{Statement}

Throughout, $x > 0$, $I_\mu$ is the modified Bessel function of the
first kind, all values $I_\mu(x)$ for $\mu \ge 0$ are positive, and
$\rho_\mu(x) = I_{\mu+1}(x)/I_\mu(x) \in (0,1)$.

\begin{theorem}[weighted Tur\'an-type unit-step monotonicity]
\label{thm:main}
For every $x > 0$ and every real $m \ge 1$,
\begin{equation}
\label{eq:main}
\varphi_m(x) \;=\;
\frac{(m-1)\,I_{m-1}(x)^2 + (m+1)\,I_{m+1}(x)^2}{m\,I_m(x)^2}
\;<\; \varphi_{m+1}(x),
\end{equation}
strictly. Equivalently, in terms of the consecutive ratios
$\rho_\mu = I_{\mu+1}/I_\mu$:
\[
\frac1m\Bigl(\frac{m-1}{\rho_{m-1}^2} + (m+1)\rho_m^2\Bigr)
\;<\;
\frac1{m+1}\Bigl(\frac{m}{\rho_m^2} + (m+2)\rho_{m+1}^2\Bigr).
\]
\end{theorem}

The weights $(m-1, m, m+1)$ are not decorative: they arise from a
specific surface expansion (Section~\ref{sec:app}), and --- as the
proof shows --- they are precisely the weights for which the
difference factorizes exactly over the classical Amos bound
\cite{Amos, Segura2011, RuizSegura}. We have not found
\eqref{eq:main} stated in the Tur\'an-type literature for modified
Bessel functions \cite{TN, Baricz1010, BariczTuranians}, though it
sits squarely in its tradition: for fixed $x$, log-concavity of
$\mu \mapsto I_\mu(x)$ (the classical Tur\'an inequality
\cite{TN, Baricz1010}) controls the \emph{first-order} behaviour of
the ratio sequence, while \eqref{eq:main} is a statement about a
\emph{weighted second-order} combination.

\section{The proof}

We use two classical facts, valid for all real orders
(\cite[10.29.1]{DLMF}): the three-term recurrence, in the two forms
\begin{equation}
\label{eq:rec}
\frac{1}{\rho_{m-1}} = \rho_m + \frac{2m}{x},
\qquad
\rho_{m+1} = \frac{1}{\rho_m} - \frac{2(m+1)}{x},
\end{equation}
and the Amos-type upper bound (Amos \cite{Amos}; proved best of its
form and valid for all real orders $\ge 0$ by Segura
\cite{Segura2011, RuizSegura}):
\begin{equation}
\label{eq:amos}
\rho_m(x) \;<\; U_m(x) := \frac{x}{a + s},
\qquad a = m + \tfrac12,\quad s = \sqrt{a^2 + x^2}.
\end{equation}

\subsection*{Step 1: elimination}

Write $u = \rho_m(x) \in (0,1)$ and $c = 1/x$. Substituting
\eqref{eq:rec} into the $\rho$-form of \eqref{eq:main} and clearing
the positive denominators $m, m+1$, the claim
$\varphi_{m+1} - \varphi_m > 0$ becomes $\Delta_m > 0$ where
\[
\Delta_m \;=\;
\frac{m^2}{u^2} - (m^2-1)\Bigl(u + 2mc\Bigr)^2
+ m(m+2)\Bigl(\frac1u - 2(m+1)c\Bigr)^2 - (m+1)^2 u^2 .
\]

\subsection*{Step 2: the exact factorization}

\begin{lemma}
\label{lem:factor}
With $S = u + 1/u$ and $P = 1/u - u$,
\[
\frac{\Delta_m}{2m(m+1)}
\;=\; (S - 3c)\,\bigl(P - (2m+1)c\bigr) \;+\; (2m+1)\,c^2 .
\]
\end{lemma}

\begin{proof}
Expanding $\Delta_m$ and collecting, the coefficients of $1/u^2$ and
$u^2$ are $\pm 2m(m+1)$, the linear-in-$c$ terms are
$-4m(m+1)c\,[(m-1)u + (m+2)/u]$, and the $c^2$ terms sum to
$8m(m+1)(2m+1)c^2$; dividing by $2m(m+1)$,
\[
\frac{\Delta_m}{2m(m+1)} = \Bigl(\frac{1}{u^2} - u^2\Bigr)
- \frac{2}{x}\Bigl((m-1)u + \frac{m+2}{u}\Bigr)
+ \frac{4(2m+1)}{x^2}.
\]
Now $(m-1)u + (m+2)/u = \tfrac12[(2m+1)S + 3P]$ and
$1/u^2 - u^2 = PS$, so the right side equals
$PS - c(2m+1)S - 3cP + 4(2m+1)c^2
= (S-3c)(P-(2m+1)c) + (2m+1)c^2$.
\end{proof}

\subsection*{Step 3: positivity of both terms}

The second factor is strictly positive: $P - (2m+1)c > 0$ is
\emph{exactly} the unit-step order-monotonicity inequality
$1/\rho_m - \rho_m > (2m+1)/x$ of the companion note
\cite{Companion}, which follows from \eqref{eq:amos} by the exact
calibration $1/U_m - U_m = (2a)/x = (2m+1)/x$ together with the
strict decrease of $t \mapsto 1/t - t$ on $(0,\infty)$.

The first factor is strictly positive: since $s > a$ in
\eqref{eq:amos}, $U_m < x/(2a) = x/(2m+1)$, so
$u < x/(2m+1) \le x/3$ for $m \ge 1$, hence
$S > 1/u > (2m+1)/x \ge 3/x = 3c$.

The third term $(2m+1)c^2$ is positive. This proves
Theorem~\ref{thm:main}. \qed

\begin{remark}[the calibration, for the third time]
\label{rem:third}
The bound $U_m$ of \eqref{eq:amos} is the positive root of
$1/t - t = (2m+1)/x$: it is the extremal function of the threshold
that appears in the factorization. In the companion note
\cite{Companion} this exact calibration was the engine of the
unit-step monotonicity of $(\log I_\nu)'$ and was shown to be
\emph{equivalent} to the Amos bound; here the same calibrated
threshold reappears, multiplied against a second factor whose
positivity is again a consequence of the same bound. Within the
audit programme from which these notes emerge, this is the third
lock opened by the same key, which suggests the calibration
identity, rather than any particular inequality, is the structural
invariant of this family of problems.
\end{remark}

\section{Application: ordering of the $\pi$-local surface expansion}
\label{sec:app}

The following objects arise in an audit-first research programme on
the two-dimensional $SU(2)$ lattice gauge theory with Wilson action
(Griffiths-II positivity on $2$-complexes; see the programme records
\cite{MasterMap} for context). For $\beta > 0$ let
$r_k = I_{k+1}(\beta)/I_1(\beta)$, $k \ge 0$, and define the
coefficient arrays ($k \ge 1$)
\[
A_k = r_{k-1}^2 \!\!\sum_{j \in \{k-2,\,k\},\; j \ge 0}\!\! (j+1)\, r_j^2,
\qquad
B_k = k\, r_{k-1}^4 .
\]
A direct computation shows $A_k/B_k = \varphi_k(\beta)$ for every
$k \ge 1$ --- including $k = 1$, where the boundary truncation of the
sum (the $j = -1$ term is absent) is annihilated by the weight
$(k-1)$ in $\varphi_k$, so no separate edge case is needed.

\begin{corollary}[determinant ordering]
\label{cor:order}
For all $\beta > 0$ and all integers $1 \le m < n$,
\[
c_{mn} \;=\; A_m B_n - A_n B_m \;=\;
B_m B_n\,\bigl(\varphi_m(\beta) - \varphi_n(\beta)\bigr) \;<\; 0 .
\]
\end{corollary}

\begin{proof}
$B_k > 0$, and $\varphi_m < \varphi_n$ by Theorem~\ref{thm:main}
iterated over unit steps from $m$ to $n$.
\end{proof}

\begin{problem}[what remains open, stated honestly]
\label{prob:open}
The positivity lemma that Corollary~\ref{cor:order} serves is: for
$t = \pi - \varepsilon$, $\varepsilon \in (0, 0.4]$, and all $\beta$
in the coupling window of the programme,
\[
W(t) \;=\; \sum_{1 \le m < n} c_{mn}\,
\bigl[(m-n)\sin((m+n)t) + (m+n)\sin((n-m)t)\bigr] \;<\; 0 .
\]
Writing $p = n-m$, $q = n+m$, the substitution $t = \pi -
\varepsilon$ gives the \emph{exact identity}, valid for every pair,
\[
W = \sum_{m<n} |c_{mn}|\,(-1)^{m+n}
\bigl[q\sin(p\varepsilon) - p\sin(q\varepsilon)\bigr],
\]
where Corollary~\ref{cor:order} supplied $c_{mn} = -|c_{mn}|$. For
the pairs with $q\varepsilon \le \pi$ the bracket is strictly
positive (decrease of $\sin(y)/y$ on $(0,\pi]$); for
$q\varepsilon > \pi$ it is not sign-definite, so the sum is a
signed-weight competition in which --- numerically --- the
odd-parity classes prevail at every $(\beta, \varepsilon)$ tested.
The claim $W<0$ is thus an \emph{alternating-parity dominance} over
the bracket weights, exact in the rewriting and open in the
dominance. High-precision
numerics (mpmath, up to $100$ significant digits) show the two parity
classes cancel to relative accuracy $\approx e^{-c\beta}$ with
$c \approx 2.03$--$2.17$ --- e.g.\ true totals
$W \approx -2.9\cdot10^{-8}$ at $\beta = 8$,
$-5.0\cdot10^{-29}$ at $\beta = 32$,
$-5.1\cdot10^{-95}$ at $\beta = 100$ (always negative) --- so any
proof must be structural rather than estimate-based, and any
numerical exploration below working precision
$\gtrsim 2.2\beta + 20$ digits sees pure rounding noise. We record
the phenomenology and the reduction here and make \emph{no claim}:
the problem is open.
\end{problem}

\section{Machine verification}

The full chain of Section~2 is machine-checked in Lean~4/Mathlib
(file \texttt{PhiMonotone.lean} accompanying this note; checked with
\texttt{lake env lean} against Lean \texttt{4.30.0-rc2}, Mathlib
commit \texttt{cd3b69b}): \texttt{amos\_calibration'} (the exact
calibration), \texttt{amos\_small} ($u < x/(2m+1)$ from
\eqref{eq:amos}), \texttt{phi\_unit\_step}
(Lemma~\ref{lem:factor} plus positivity), and
\texttt{phi\_step\_of\_recurrences} (Theorem~\ref{thm:main} with the
recurrences \eqref{eq:rec} and the bound \eqref{eq:amos} entering as
the named hypotheses \texttt{hrec1}, \texttt{hrec2},
\texttt{hAmos}). The axiom oracle returns Lean's three standard
axioms for all four theorems:
\begin{verbatim}
'amos_calibration'' depends on axioms:
    [propext, Classical.choice, Quot.sound]
'amos_small' depends on axioms:
    [propext, Classical.choice, Quot.sound]
'phi_unit_step' depends on axioms:
    [propext, Classical.choice, Quot.sound]
'phi_step_of_recurrences' depends on axioms:
    [propext, Classical.choice, Quot.sound]
\end{verbatim}
(exit code $0$; no \texttt{sorry}, no \texttt{admit}). As in the
companion note, what is machine-checked is the ordered-field
implication [recurrences $+$ Amos bound $\Rightarrow$ monotonicity];
the analytic inputs themselves are classical and cited.

\section{Numerical audit}

The factorization of Lemma~\ref{lem:factor} was verified symbolically
(sympy: residual $0$). The reduction $\varphi_k = A_k/B_k$ (including
the $k=1$ edge), the identity
$\Delta_m = m(m+1)(\varphi_{m+1}-\varphi_m)$, and the positivity of
both factors were verified in mpmath at $40$ significant digits on
$\beta \in \{0.3, 2, 8, 32, 100, 1000\}$, $m \le 11$; the unit-step
increase of $\varphi_m$ for \emph{real} $m$ was checked on
$m \in \{1, 1.3, 2, 2.71, 5.5, 9, 14.25, 40\}$,
$x \in \{0.2, 1, 3.7, 15, 80, 400\}$ (minimal margin
$7.5\cdot10^{-5}$ at $x=400$, $m=1$); and the strict increase of the
integer sequence was verified for $k \le 200$ at
$\beta \in \{2, 8, 32, 100\}$. Scripts and sources are archived in
the programme repository:
\begin{center}\small
\url{https://github.com/lluiseriksson/THE-ERIKSSON-PROGRAMME/tree/main/surface-theorem}
\end{center}

\subsection*{Acknowledgements}
This note is part of an audit-first research programme whose records,
Lean sources and numerical suites are public at
\url{https://github.com/lluiseriksson/THE-ERIKSSON-PROGRAMME}. The
statement is presented as a lemma with an application and an open
problem, in that order; nothing here is claimed about the open
problem beyond its precise reduction, and nothing is claimed about
four-dimensional gauge theory.

\begin{thebibliography}{10}\small

\bibitem{Amos}
D.~E. Amos,
\emph{Computation of modified Bessel functions and their ratios},
Math. Comp. \textbf{28} (1974), 239--251.

\bibitem{Segura2011}
J.~Segura,
\emph{Bounds for ratios of modified Bessel functions and associated
Tur\'an-type inequalities},
J. Math. Anal. Appl. \textbf{374} (2011), 516--528.

\bibitem{RuizSegura}
D.~Ruiz-Antol\'in and J.~Segura,
\emph{A new type of sharp bounds for ratios of modified Bessel
functions},
J. Math. Anal. Appl. \textbf{443} (2016), 1232--1246;
arXiv:1606.02008.

\bibitem{TN}
V.~R. Thiruvenkatachar and T.~S. Nanjundiah,
\emph{Inequalities concerning Bessel functions and orthogonal
polynomials},
Proc. Indian Acad. Sci. Sect. A \textbf{33} (1951), 373--384.

\bibitem{Baricz1010}
\'A.~Baricz,
\emph{On Tur\'an type inequalities for modified Bessel functions},
Proc. Amer. Math. Soc. \textbf{141} (2013), 523--532;
arXiv:1010.3346.

\bibitem{BariczTuranians}
\'A.~Baricz and T.~K. Pog\'any,
\emph{Bounds for Tur\'anians of modified Bessel functions},
Expo. Math. \textbf{33} (2015), 223--251; arXiv:1202.4853.

\bibitem{DLMF}
F.~W.~J. Olver et al.,
\emph{NIST Digital Library of Mathematical Functions},
\S10.29, \url{https://dlmf.nist.gov/10.29}.

\bibitem{Companion}
L.~Eriksson,
\emph{An elementary recurrence--Amos proof of the unit-step
order-monotonicity of $(\log I_\nu)'$, with a Feynman--Hellmann
application to two-dimensional lattice gauge theory},
July 2026;
\url{https://github.com/lluiseriksson/THE-ERIKSSON-PROGRAMME/tree/main/papers/bessel-amos-fh}.

\bibitem{MasterMap}
L.~Eriksson,
\emph{THE MASTER MAP: an audit-first navigation guide to the
conditional construction of 4D $SU(N)$ Yang--Mills with mass gap},
ai.viXra:2602.0096, and the programme repository
\url{https://github.com/lluiseriksson/THE-ERIKSSON-PROGRAMME}.

\end{thebibliography}

\end{document}
