\documentclass[11pt]{article}
\usepackage{cmap}
\usepackage[margin=1.1in]{geometry}
\usepackage{amsmath,amssymb,amsthm}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\input{glyphtounicode}
\pdfgentounicode=1
\usepackage{xcolor}
\usepackage{booktabs}
\usepackage{hyperref}
\definecolor{linkblue}{rgb}{0.1,0.2,0.55}
\hypersetup{colorlinks=true,linkcolor=linkblue,citecolor=linkblue,urlcolor=linkblue,
pdftitle={Global Ratio Monotonicity for a Killed von Mises Bridge},
pdfauthor={Lluis Eriksson}}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{conjecture}[theorem]{Conjecture}
\theoremstyle{remark}
\newtheorem{remark}[theorem]{Remark}
\newtheorem{convention}[theorem]{Convention}

\newcommand{\FA}{F_{\!A}}
\newcommand{\FB}{F_{\!B}}
\newcommand{\dd}{\,d}

\title{Global Ratio Monotonicity for a Killed von Mises Bridge\\
\large [COMPLETE-THEOREM MANUSCRIPT --- SEAL IN PROGRESS; DO NOT SUBMIT.\\
Pending relay lemmas are marked {[SLOT]}.]}
\author{Lluis Eriksson}
\date{July 2026}

\begin{document}
\maketitle

\begin{abstract}
For $\beta>0$ let $I_m=I_m(\beta)$ denote modified Bessel functions of the
first kind, and set $a_m = I_m^2\,[(m-1)I_{m-1}^2+(m+1)I_{m+1}^2]$,
$b_m = m\,I_m^4$, $\FA(t)=\sum_{m\ge1}a_m\sin(mt)$,
$\FB(t)=\sum_{m\ge1}b_m\sin(mt)$, and $E=\FA/(2\FB)$. The global
ratio-monotonicity problem for the surface expansion of a two-dimensional
$SU(2)$ lattice gauge observable is: (i) $\FB>0$ on $(0,\pi)$, and (ii)
$E'<0$ on $(0,\pi)$, for every $\beta>0$. We prove (i) twice, and prove (ii) for $0<\beta\le3$ via an exact
pair-identity skeleton plus an archived interval certificate executed by
two independent interval-arithmetic implementations. The
structural core is exact: $E$ is, as an algebraic identity, the mean of
$\cos\psi$ under the midpoint law of a four-step killed von Mises
bridge; the generating kernels reduce, via the Neumann addition theorem,
to two-dimensional integrals of a \emph{single} Bessel function whose
saddle deficit is an exact sum of two squares; and exact saddle
cancellations yield the coefficients of the (verified) closed
second-order law $E=\cos(t/2)(1-c(t)/\beta)+O(\beta^{-2})$,
$c(t)=(4\cos^2(t/4)-1)/(2\cos(t/4)\cos(t/2))$. Three certified negative
results (interval arithmetic, two implementations, nested enclosures)
kill every monotone full-path coupling, with an exact mechanism at
threshold $\beta|\cos t|=3/2$. For the remaining range we give a
two-scale closure map with a verified candidate threshold $\beta_0=15$
and, at the $\pi$-endpoint, exact identities for the cubic coefficient
$c_3$ (telescoped alternating form, integral form, parity) together with
its verified prefactor law. Every claim is labelled exact / certified / verified; the
machine-checked lemmas are Lean~4/Mathlib, machine-checked modulo
classical Bessel inputs carried as named hypotheses.
\end{abstract}

\section{Introduction}

This paper consolidates the resolution status of a two-part positivity
problem that arises as the last analytic gap in a surface (character)
expansion for two-dimensional $SU(2)$ lattice gauge theory
\cite{Companion,PhiLemma,WronskianNote}. The problem is stated entirely in
terms of modified Bessel functions and trigonometric series, and we treat
it as such; no claim of physical consequence beyond the stated expansion
is made, and importance is not inherited by thematic proximity.

Throughout, $\beta>0$, $I_m=I_m(\beta)$, and
\begin{equation}\label{eq:coeffs}
a_m = I_m^2\bigl[(m-1)I_{m-1}^2+(m+1)I_{m+1}^2\bigr],\qquad
b_m = m\,I_m^4,\qquad m\ge1 .
\end{equation}
Both sequences decay super-exponentially, so
\begin{equation}\label{eq:FAFB}
\FA(t)=\sum_{m\ge1}a_m\sin(mt),\qquad
\FB(t)=\sum_{m\ge1}b_m\sin(mt),\qquad
E(t)=\frac{\FA(t)}{2\FB(t)},
\end{equation}
are real-analytic on $\mathbb{R}$, and the Wronskian
$W=2(\FA'\FB-\FA\FB')$ satisfies $W=4\FB^2\,(E)'$ wherever $\FB\neq0$
\cite{WronskianNote}.

\begin{conjecture}[global ratio monotonicity]\label{conj:main}
For every $\beta>0$: \textup{(i)} $\FB>0$ on $(0,\pi)$;
\textup{(ii)} $E'<0$ on $(0,\pi)$.
\end{conjecture}

Part (i) is proved below (twice). Part (ii) is proved for $0<\beta\le3$
(Theorem~\ref{thm:B}; exact skeleton plus the archived certificate
\texttt{certify\_thmB.py} and its Arb twin), is supported by an exact
structural reduction and a verified truth map for $\beta\ge15$, and is
otherwise open; Section~\ref{sec:closure} states the precise two-scale
closure map. The conjecture was born inside this programme (it is the
global form of an $\varepsilon$-window positivity required by the surface
expansion); no external pedigree is claimed.

\begin{convention}[tricotomy]\label{conv:tri}
Every assertion below carries one of three labels. \emph{Exact}: a
proved identity or inequality. \emph{Certified}: a numerical statement
rigorously enclosed by interval arithmetic with outward rounding,
together with complete analytic bounds for domains and truncations.
Precision-stability re-runs and independent implementations are
\emph{software audits}, recorded separately with their declared
coverage: the certified negatives
\texttt{certify\_time3\_negative} and \texttt{certify\_bridge\_matrix}
carry a \emph{full twin audit} (nested re-computation plus a complete
independent Arb implementation, both transcripts archived); for
\texttt{certify\_thmB} both implementations exist and the Arb
transcript is archived --- its row is \emph{amber} only because the
\texttt{mpmath.iv} canonical transcript is still in progress; the
$[6,15]$ coverage will carry a \emph{planned pre-registered sampled
independent audit} (Remark~\ref{rem:certaudit}). \emph{Verified}: high-precision
numerics with stated grids and precision policies, not a proof. No
verified statement is used as a hypothesis of an exact one.
\end{convention}

\section{The theorem and the relay}\label{sec:relay}

\begin{theorem}[the Surface Theorem; \textbf{[SLOT: seal completion]}]
\label{thm:main}
For every $\beta>0$: \textup{(i)} $\FB>0$ on $(0,\pi)$, and
\textup{(ii)} $E'<0$ on $(0,\pi)$.
\end{theorem}

Part (i) is Theorem~\ref{thm:A}. Part (ii) is proved by the following
relay. Rows marked \textbf{[SLOT]} are pending; rows marked \emph{amber}
have one archived witness with the second in progress; all other rows
are complete with the stated witnesses.

\begin{center}\footnotesize
\begin{tabular}{@{}lllll@{}}
\toprule
range & method & artifact & transcript & state\\
\midrule
$(0,\tfrac1{20}]$ & small-$\beta$ lemma & Lem.~\ref{lem:smallbeta} & --- & complete\\
$[\tfrac1{20},3]$, all $t$ & pair skeleton & \texttt{certify\_thmB} & Arb & \emph{amber}\\
$[3,6]$, $t{\in}[0.6,\pi{-}\tfrac{1.5}{\beta}]$ & series intervals & \texttt{certify\_bulk} & Arb & \emph{amber}\\
$[6,15]$, $t{\in}[0.6,\pi{-}\tfrac{1.5}{\beta}]$ & exp-factored integrator & \texttt{exp\_integrator} & \textbf{[SLOT]} & local box: Prop.~\ref{prop:localbox}\\
$[15,\infty)$, bulk & analytic relay & \textbf{[SLOT]} & --- & pending\\
$t\in(0,0.6]$, $\beta\ge3$ & sinc + minor criterion & \textbf{[SLOT: sinc-cert]} & --- & design ratified\\
$t\in[\pi{-}\tfrac{C}{\beta},\pi)$ & $c_3$ + radius & \textbf{[SLOT: 4 const.]} & --- & $A$ verified\\
\bottomrule
\end{tabular}
\end{center}

Splice boundaries are explicit and overlapping: $\beta=3$, $6$, $15$;
$t=0.6$ and $t=\pi-C/\beta$, with $C_{\rm WIN}=1.5$ a candidate pending
$C_{\rm mirror}\le1.5\le C_{\rm quad}$. The remainder of this manuscript
is the completed material, verbatim from the consolidation paper.

\begin{remark}[supersession of the $e_4$ route]\label{rem:e4dead}
An earlier plan closed the block $t\in(0,0.6]$, $\beta\ge3$ by an
$e_4$-coefficient lemma with a fixed radius; it is superseded, not
silently deleted. A feasibility check showed the sinc criterion below
dominating with ratios from $15$ up to $10^{80}$ across
$\beta\in[3,15]$ for $t\le0.3$ (\emph{verified}, design-grade), while
the apparent infeasibility of $(0,0.6]$ belonged to the cancellation
of the series tool, which the exp-factored integrator does not
suffer. The ratified route has three components: the exact sinc
positivity (Lemma~\ref{lem:sinc}), cheap one-dimensional-in-$\beta$
minor-criterion certificates on $(0,t_*(\beta)]$, and the extension
of the integrator strip from $t=0.6$ down to $t_*(\beta)\approx0.45$
with splice overlap. The slot's pending content is the full minor
criterion with its uniform overlap; the $e_4$ lemma is retired.
\end{remark}

\begin{lemma}[sinc positivity]\label{lem:sinc}
Let $1\le m<n$ and $T_{mn}(t)=(m-n)\sin((m+n)t)+(m+n)\sin((n-m)t)$
as in Section~\textup{4}. Then $T_{mn}(t)>0$ for
$0<t\le\pi/(m+n)$.
\end{lemma}

\begin{proof}
Write $s=m+n$, $p=n-m\ge1$, so $T_{mn}(t)=s\sin(pt)-p\sin(st)$.
Since $\sin(x)/x$ is strictly decreasing on $(0,\pi]$ and
$0<pt<st\le\pi$, we get $\sin(pt)/(pt)>\sin(st)/(st)$, i.e.\
$st\sin(pt)>pt\sin(st)$; dividing by $t>0$ gives
$s\sin(pt)>p\sin(st)$.
\end{proof}

\begin{lemma}[{relay criterion for $\beta\in[6,15]$}]
\label{lem:relaycrit}
Let $\langle f\rangle=\iint_{[-\pi,\pi]^2}Kf\,ds\,d\alpha$ with
$K=I_1(2\beta R)/R>0$, and let $\bar E$ be any rational constant. Then
$E'<0$ at $(t,\beta)$ if and only if $\langle D\rangle>0$ and
\[
W_c:=\langle N_t\rangle\langle D\rangle
+\langle (N-\bar E D)\,\partial_t\!\log K\rangle\langle D\rangle
-\bigl(\langle N\rangle-\bar E\langle D\rangle\bigr)
\langle D\,\partial_t\!\log K\rangle<0
\]
(the $\bar E$-terms cancel identically). The certificate encloses the
five integrals by exp-factored interval cells: $z=2\beta R$ is
linearized at the exact rational cell center with interval-gradient
remainder; the linear exponential integrates exactly; and the two
scaled Bessel factors are
$A(z)=e^{-z}I_1(z)/z$ --- strictly decreasing, since
$(\log A)'=I_0/I_1-1-2/z<0$ follows from the classical lower bound
at $\nu=0$ (Amos \cite{Amos}; \cite{Segura2011}, Thm.~1) and
$\sqrt{1+z^2}\le z+1$ ---
enclosed by pure endpoint evaluation, and $B(z)=e^{-z}H(z)$, enclosed
unconditionally by $[e^{-z_{\rm hi}}H(z_{\rm lo}),
e^{-z_{\rm lo}}H(z_{\rm hi})]$; the identity
$K\,\partial_t\!\log K=32\beta^3\cos(t/4)\,(\cos(t/4))'\,(1-P-Q)H(z)$
removes $I_1(z)/z$ from the covariance integrands exactly. The
$(t,\beta)$-box layer is Proposition~\ref{prop:localbox}; the global
$[6,15]$ coverage remains the pending content of the relay row.
\end{lemma}

\begin{proposition}[{certified; local box:
$(t,\beta)\in[1.5,1.51]\times[8,8.05]$}]
\label{prop:localbox}
For every $(t,\beta)\in[1.5,1.51]\times[8,8.05]$ one has
$\langle D\rangle>0$ and $W_c<0$; hence, by
Lemma~\ref{lem:relaycrit}, $E'(t,\beta)<0$ on this box.
\end{proposition}

\begin{proof}[Certificate]
Ball-arithmetic machine certificate (python-flint 0.9.0 / Arb).
The executed script is identified by SHA-256 prefix
\texttt{1d888e99} and is archived byte-exact in the repository,
together with the self-contained transcript (both file names carry
the hash; the transcript's provenance header records the full hash,
library versions, and stage parameters). Three stages, one
transcript. \emph{Point} $(1.5,8)$: $1{,}951{,}141$ cells,
$W_c/\langle D\rangle^2\subset[-0.4546,-0.1752]$.
\emph{Stability} (precision $120$, $\Delta z\le 0.12$):
$2{,}500{,}336$ cells, ratio $\subset[-0.4304,-0.1978]$, nested
inside the point enclosure. \emph{Box}: union of nine parametric
sub-boxes ($\Delta\beta\le0.02$, $\Delta t\le0.005$), each with
strictly negative $W_c$ enclosure and $\langle D\rangle>0$ (worst
$W_c$ upper end $\le-2.8\times10^{21}$), $22{,}986{,}387$ cells in
total, $7558$~s. All comparisons are strict ball comparisons; the
displayed intervals are outward-rounded supersets of the printed
enclosures. This proposition validates the mechanism and calibrates
the coverage design; it does \emph{not} close the $[6,15]$ row,
which stays open until the full region $6\le\beta\le15$,
$0.6\le t\le\pi-1.5/\beta$ is covered.
\end{proof}

\begin{remark}[what certifies, and what audits]
\label{rem:certaudit}
The mathematical certificate for each sub-box is its \emph{single}
rigorous Arb enclosure: outward-rounded ball arithmetic makes one
honest enclosure a proof of the strict inequalities, per sub-box.
The precision-$120$ stability stage (with its nesting inside the
point enclosure) and the \emph{planned} pre-registered sampled
\texttt{mpmath.iv} reproduction are \emph{implementation audits}:
they test the
software that produced the enclosures, not the inequalities
themselves, and they are declared as such in the relay table. This
fixes the convention for the $[6,15]$ coverage: one certificate-grade
Arb enclosure per sub-box; stability and sampling audit the pipeline.
\end{remark}

\begin{lemma}[safe square roots in the archived run]
\label{lem:dormant}
In the archived script (\textup{SHA-256} \texttt{1d888e99}), the
square-root guard clamps a negative lower endpoint to $0$ and, when
the \emph{upper} endpoint ball straddles $0$, returns the dyadic
upper bound $2^{8-p/2}$ at working precision $p\ge90$. This bound is
valid throughout the run. Indeed, the guarded inputs are true
squares: $R^2=4[c_0^2(1-P-Q)+PQ]$ with $c_0^2\le1$, $P,Q\in[0,1]$,
and $z^2=4\beta^2R^2$ with $\beta\le15$, so $|R^2|\le16$,
$|z^2|\le14400$; the corresponding \emph{balls} are built by at most
twelve correctly rounded operations on factors of unit size, hence
$|R^2_{\mathrm{ball}}|<16.01$ and $|z^2_{\mathrm{ball}}|<14401<2^{14}$
(magnitude means $|\mathrm{mid}|+\mathrm{rad}$). The upper-endpoint
ball $u=\mathrm{mid}+\mathrm{rad}$ involves two exact arf-to-ball
conversions and one correctly rounded addition, so
$\mathrm{rad}(u)\le\mathrm{ulp}(2^{14})=2^{15-p}$. If $u$ straddles
$0$ then $|\mathrm{mid}(u)|\le\mathrm{rad}(u)$, hence the true upper
endpoint is at most $2\,\mathrm{rad}(u)\le2^{16-p}$ and its square
root is at most $2^{8-p/2}$ --- exactly the coded constant. (In the
definitive script the branch is replaced by
\texttt{abs\_upper()}, an automatic outer bound with no ad-hoc
constant; the present lemma covers the archived run.)
\end{lemma}

\subsection*{The two scaled functions}

The certificate's precision rests on three one-line facts. \emph{(a)}
$A(z)=e^{-z}I_1(z)/z$ is strictly decreasing on $(0,\infty)$: by the
classical lower bound $I_{\nu+1}(z)/I_\nu(z)\ge
z/\bigl(\nu+1+\sqrt{(\nu+1)^2+z^2}\bigr)$, valid for $z>0$ and
$\nu\ge0$ (Amos \cite{Amos}; both members of the classical two-sided
pair are proved in \cite{Segura2011}, Theorem~1), taken at $\nu=0$,
together with $\sqrt{1+z^2}\le z+1$ (square both sides),
$I_1/I_0>z/(z+2)$, which is
exactly $(\log A)'=I_0/I_1-1-2/z<0$; hence $A$ is enclosed over any
$z$-interval by its two endpoint values. \emph{(b)} The kernel--score
product collapses algebraically,
$K\,\partial_t\!\log K=32\beta^3\cos(t/4)(\cos(t/4))'(1-P-Q)\,H(z)$
(the factor $I_1(z)/z$ cancels between $K$ and the denominator of the
regular score), so the covariance integrands involve only
$B(z)=e^{-z}H(z)$. \emph{(c)} $B$ is enclosed \emph{unconditionally} by
$[e^{-z_{\rm hi}}H(z_{\rm lo}),\,e^{-z_{\rm lo}}H(z_{\rm hi})]$, since
$e^{-z}$ decreases and $H$ (a positive-coefficient entire series)
increases; after centering, the $B$-terms are subdominant and this
crude-but-safe bound suffices. Together with the exact linear-exponential
cell integrals, every source of exponential interval growth is either
integrated analytically or carried by a monotone one-dimensional
profile --- the certificate is structure, not brute force.



\section{The spectral framework}\label{sec:frame}

\begin{convention}[normalizations]\label{conv:norm}
On $(0,\pi)$ define the kernel, entry law, and smoothing operator
\begin{align}
q_1(u,v) &= 2\,e^{\beta\cos u\cos v}\sinh(\beta\sin u\sin v)
          \;=\; 4\sum_{m\ge1}I_m\sin(mu)\sin(mv), \label{eq:q1}\\
s_1(u)   &= \tfrac{\beta}{2}\sin u\,e^{\beta\cos u}
          \;=\; \sum_{m\ge1}m\,I_m\sin(mu), \label{eq:s1}\\
(Qf)(v)  &= \frac{1}{2\pi}\int_0^\pi q_1(v,u)f(u)\dd u ,\label{eq:Q}
\end{align}
so that $Q\sin(m\cdot)=I_m\sin(m\cdot)$. Both closed forms are classical
consequences of the generating function
$e^{\beta\cos\theta}=I_0+2\sum_{m\ge1}I_m\cos(m\theta)$; \eqref{eq:s1}
follows by differentiation. All ten verification artifacts in the
repository use exactly these normalizations.
\end{convention}

Since $q_1>0$ on $(0,\pi)^2$ and $s_1>0$ on $(0,\pi)$, the walk with step
kernel $q_1$ killed on $\{0,\pi\}$ is well defined; $H(\psi)=\beta
\sin(\psi/2)I_1(2\beta\cos(\psi/2))$ and $K_2(\varphi)=I_0(2\beta
\cos(\varphi/2))$ appear below. Note $\FB=Q^3s_1$ and, by the sine
orthogonality computation of Lemma~\ref{lem:bridge}, $\FA$ is the
companion cosine-weighted integral.

\section{Theorem A: positivity of $\FB$}\label{sec:thmA}

\begin{theorem}[exact]\label{thm:A}
$\FB(t)>0$ for all $t\in(0,\pi)$ and all $\beta>0$.
\end{theorem}

\begin{proof}[First proof (kernel positivity)]
$\FB=Q^3s_1$ by \eqref{eq:Q} and $b_m=mI_m^4$. Each application of $Q$
integrates a positive function against the strictly positive kernel
\eqref{eq:q1}; $s_1>0$ on $(0,\pi)$. Hence $\FB>0$ pointwise.
\end{proof}

\begin{proof}[Second proof (symmetric-decreasing convolution)]
With $K_2(\varphi)=I_0(2\beta\cos(\varphi/2))$ and $P_4=K_2*K_2$ (circle
convolution; Section~\ref{sec:reduction}), $\FB=-\tfrac12P_4'$. The
function $K_2$ is even and \emph{strictly decreasing in $|\varphi|$ on
$[0,\pi]$} ($I_0$ strictly increasing, $\cos(\varphi/2)$ strictly
decreasing). By the layer-cake representation, $K_2=\int_0^\infty
\mathbf 1_{A_\lambda}\dd\lambda$ with $A_\lambda$ centered arcs whose
half-lengths $a(\lambda)$ fill $(0,\pi)$ continuously (strict decrease).
For centered arcs, $(\mathbf 1_{A}*\mathbf 1_{B})(t)=|A\cap(t+B)|$ is
even and nonincreasing in $|t|$ on $[0,\pi]$, and \emph{strictly}
decreasing on $||a-b|,\min(a+b,2\pi-a-b)|$. Hence
$P_4(t)=\iint |A_\lambda\cap(t+A_\mu)|\dd\lambda\dd\mu$ is even,
nonincreasing on $[0,\pi]$, and strictly decreasing on $(0,\pi)$ because
for every $t\in(0,\pi)$ a positive measure of pairs $(\lambda,\mu)$ has
$|a(\lambda)-a(\mu)|<t<\min(a+b,2\pi-a-b)$ (the half-lengths fill
$(0,\pi)$). At the derivative level: each overlap $t\mapsto|A_\lambda\cap(t+A_\mu)|$
is Lipschitz with a.e.\ derivative $-1$ in the strict regime and $0$
otherwise (on $[0,\pi]$); since $P_4$ is analytic, dominated convergence
gives $P_4'(t)=\iint\partial_t|A_\lambda\cap(t+A_\mu)|\dd\lambda\dd\mu
\le-\,\mu\{(\lambda,\mu):\text{strict at }t\}<0$ pointwise on $(0,\pi)$.
Thus $\FB=-\tfrac12P_4'>0$ on $(0,\pi)$.
\end{proof}

\section{Theorem B: ratio monotonicity for $0<\beta\le3$}\label{sec:thmB}

\begin{lemma}[exact; Wronskian identity]\label{lem:wronskian}
For real arrays with $\sum m(|a_m|+|b_m|)<\infty$,
\begin{equation}\label{eq:wronskian}
W:=2(\FA'\FB-\FA\FB')=\sum_{1\le m<n}c_{mn}\,T_{mn}(t),\qquad
c_{mn}=a_mb_n-a_nb_m,
\end{equation}
where $T_{mn}(t)=(m-n)\sin((m+n)t)+(m+n)\sin((n-m)t)$.
\end{lemma}

\begin{proof}
Product-to-sum: $T_{mn}=2[m\cos(mt)\sin(nt)-n\sin(mt)\cos(nt)]=:2G(m,n)$,
antisymmetric with $G(n,n)=0$; since $c_{mn}$ is antisymmetric,
$\sum_{m<n}c_{mn}2G=2\sum_{m,n}a_mb_nG(m,n)=2[\FA'\FB-\FA\FB']$
\textup{(}first recorded in \cite{WronskianNote}\textup{)}.
\end{proof}

The coefficient input is the weighted Tur\'an-type lemma
(Appendix~\ref{app:phi}, self-contained): $\varphi_m=a_m/b_m$ is strictly
increasing in $m\ge1$ for every $\beta>0$, hence \emph{every minor is
negative}: $c_{mn}<0$ for $m<n$ (exact).

\begin{lemma}[exact; the pair identities and bounds]\label{lem:pairs}
For $t\in(0,\pi)$, with $p=n-m$, $q=n+m$:
\begin{itemize}
\item[(a)] $T_{12}=4\sin^3t$;\quad $T_{13}=16\sin^3t\,\cos t$;\quad
$T_{24}=8\sin^3(2t)$, so $|T_{13}|\le16\sin^3t$ and
$|T_{24}|\le64\sin^3t$.
\item[(b)] $T_{m,m+1}\ge0$ for every $m\ge1$.
\item[(c)] $|T_{mn}|\le\dfrac{\pi^3}{48}\,pq(q^2+p^2)\,\sin^3t$ for all
$m<n$.
\end{itemize}
\end{lemma}

\begin{proof}
(a) $3\sin u-\sin3u=4\sin^3u$ gives $T_{12}$ (at $u=t$) and $T_{24}$ (at
$u=2t$); $T_{13}=4\sin2t-2\sin4t=4\sin2t(1-\cos2t)=16\sin^3t\cos t$.
(b) $T_{m,m+1}=(2m+1)\sin t-\sin((2m+1)t)\ge0$ since
$|\sin(qt)|\le q\sin t$ on $(0,\pi)$ (induction on $q$).
(c) $T_{mn}$ and its first two derivatives vanish at $t=0$ and $t=\pi$
(the parities of $m\pm n$ agree), and $|T_{mn}'''|\le
pq^3+qp^3=pq(q^2+p^2)$; hence
$|T_{mn}(t)|\le\frac{pq(q^2+p^2)}{6}\min(t,\pi-t)^3$, and
$\sin t\ge\frac2\pi\min(t,\pi-t)$ gives the claim.
\end{proof}

\begin{theorem}[exact modulo the certified computation
\texttt{certify\_thmB.py}]\label{thm:B}
For $0<\beta\le3$:\; $W<0$ on $(0,\pi)$, hence $E'<0$ on $(0,\pi)$.
\end{theorem}

\begin{proof}
Drop the adjacent pairs $m\ge2$ (each contributes $c_{m,m+1}T_{m,m+1}
\le0$ by Lemma~\ref{lem:pairs}(b) and $c<0$); keep $T_{12}$, $T_{13}$,
$T_{24}$ exactly and majorize every other pair by
Lemma~\ref{lem:pairs}(c). Pointwise in $t$,
\[
\frac{W(t)}{\sin^3t}\;\le\;\mathrm{Crit}(\beta):=
-4|c_{12}|+16|c_{13}|+64|c_{24}|
+\frac{\pi^3}{48}\!\!\sum_{\substack{p\ge2\\(m,n)\ne(1,3),(2,4)}}\!\!
pq(q^2+p^2)\,|c_{mn}| .
\]
The certificate \texttt{certify\_thmB.py} (Section~\ref{sec:verif})
encloses $\mathrm{Crit}$ in interval arithmetic over adaptive
$\beta$-boxes covering $[1/20,3]$ ($I_m$ increasing in $\beta$ gives
rigorous boxes; cutoff $M=60$ with the tail itself computed in
intervals and closed by a derived, per-box-asserted geometric ratio) and
certifies $\mathrm{Crit}<0$ on all of $[1/20,3]$; the
script then re-certifies \emph{every} box at precision $+70$ before
reporting success (the box count is invocation- and implementation-dependent; the
independent Arb implementation \texttt{certify\_thmB\_arb.py} certifies
the same coverage with $86$ boxes and a full stability pass, transcript
archived). For $0<\beta\le1/20$, Lemma~\ref{lem:smallbeta} applies.
\end{proof}

\begin{lemma}[exact; small $\beta$]\label{lem:smallbeta}
For $0<\beta\le1/20$:\; $\mathrm{Crit}(\beta)<0$.
\end{lemma}

\begin{proof}
Write $\kappa=\beta/2\le1/40$, $L_m=\kappa^m/m!$,
$U_m=e^*L_m$ with $e^*=e^{\beta^2/4}\le e^{1/1600}<1.001$; then
$L_m\le I_m\le U_m$ (positive series). \emph{Leading minor:}
$a_2b_1\ge(I_1^2I_2^2)(I_1^4)\ge L_1^6L_2^2=\kappa^{10}/4$ and
$a_1b_2=(2I_1^2I_2^2)(2I_2^4)\le4U_1^2U_2^6=(e^*)^8\kappa^{14}/16$,
so
\[
|c_{12}|\ \ge\ \frac{\kappa^{10}}{4}
\Bigl(1-\frac{(e^*)^8\kappa^4}{4}\Bigr)
\ \ge\ \frac{\kappa^{10}}{4}\,(1-10^{-6}).
\]
\emph{The three visible positive terms:}
$a_3\le I_3^2\cdot(2I_2^2+4I_4^2)\le(e^*)^4\kappa^{10}/24\cdot(1+10^{-4})$
and $b_1\le(e^*)^4\kappa^4$ give
$16|c_{13}|\le16(a_3b_1+a_1b_3)\le0.70\,\kappa^{14}$;
$64|c_{24}|\le64(a_4b_2+a_2b_4)\le10^{-3}\kappa^{14}$ (both products
have $\kappa$-degree $\ge22$); the largest remaining weighted pair is
$(1,4)$: $\frac{\pi^3}{48}\cdot510\cdot(a_4b_1+a_1b_4)
\le0.2\,\kappa^{18}\le10^{-6}\kappa^{14}$.
\emph{All other pairs} (those with $s:=m+n\ge6$, $p\ge2$): using
$a_nb_m+a_mb_n\le2mn(e^*)^8\kappa^{4s-2}
[((m-1)!m!)^{-2}n!^{-4}+((n-1)!n!)^{-2}m!^{-4}]
\le s^2(e^*)^8\kappa^{4s-2}$ (the factorial bracket is $\le2$ and
$2mn\le s^2/2$), $pq(q^2+p^2)\le 2s^4$, and at most $s$ pairs per level,
\[
\frac{\pi^3}{48}\!\!\sum_{s\ge6}\!\!\;(\text{level }s)
\le\frac{\pi^3}{48}(e^*)^8\sum_{s\ge6}2s^7\kappa^{4s-2}
\le0.66\,\kappa^{22}\cdot6^7\sum_{j\ge0}\bigl[(7/6)^7\kappa^4\bigr]^j
\le3.7\times10^5\,\kappa^{22}\le10^{-7}\kappa^{14},
\]
where the level ratio $s\mapsto s+1$ is at most
$(7/6)^7\kappa^4<10^{-5}<\tfrac12$ (so the geometric closure is valid),
and $\kappa^{8}\le40^{-8}$ absorbs the constant. Collecting:
\[
\mathrm{Crit}\le-\kappa^{10}(1-10^{-6})+0.71\,\kappa^{14}
=-\kappa^{10}\bigl(1-10^{-6}-0.71\,\kappa^4\bigr)<0,
\]
since $\kappa^4\le40^{-4}<4\cdot10^{-7}$. (All numbered constants above
are crude upper counts; each step loses factors, none gains.)
\end{proof}

\begin{remark}[verified; the sharp family and saturation]
\label{rem:saturation}
The sharper per-pair bound $|T_{mn}|\le\frac{pq(q^2-p^2)}{6}\sin^3t$ and
the adjacent floor $T_{m,m+1}\ge2m\sin^3t$ \textup{(}note the sign: the
floor is a \emph{lower} bound on a positive quantity\textup{)} are
verified far beyond the range used ($q\le151$; $13$ pairs to $(10,30)$,
max ratio $1.0$) and would extend Theorem~\ref{thm:B} to
$\beta\approx3.5$; their per-pair certification is a scheduled upgrade.
Beyond that the decomposition is \emph{saturated}: with the exact far
block (cancellation included) the criterion fails at $\beta=6$ by
$+2.2\times10^{14}$ while the true $W/\sin^3t$ remains negative
($-6.9\times10^{12}$) --- a verified adversarial finding about the
method, and the reason the large-$\beta$ closure uses different
machinery (Sections~\ref{sec:reduction}--\ref{sec:closure}).
\end{remark}

\section{The bridge structure of $E$ (exact)}\label{sec:bridge}

Let $k=q_1\!\circ\!q_1$ denote the two-step kernel,
$k(\psi,t)=4\sum_{m\ge1}I_m^2\sin(m\psi)\sin(mt)$, and define
\begin{equation}\label{eq:omega}
\omega_t(\dd\psi)\;\propto\;(Qs_1)(\psi)\,k(\psi,t)\dd\psi
\qquad\text{on }(0,\pi),
\end{equation}
the law of the time-$2$ position (midpoint) of the four-step bridge
entering from $0$ with law $s_1$ and conditioned to end at $t$.

\begin{lemma}[exact]\label{lem:bridge}
$\displaystyle \int_0^\pi (Qs_1)\,k(\cdot,t)\dd\psi=2\pi\FB(t)$ and
$\displaystyle \int_0^\pi \cos\psi\,(Qs_1)\,k(\cdot,t)\dd\psi=\pi\FA(t)$.
Hence, as an identity,
\[
E(t)\;=\;\mathbb{E}_{\omega_t}[\cos\psi]\;=\;\frac{\FA(t)}{2\FB(t)} .
\]
\end{lemma}

\begin{proof}
$(Qs_1)(\psi)=\sum m I_m^2\sin(m\psi)$. By sine orthogonality the first
integral picks the diagonal, giving $2\pi\sum mI_m^4\sin(mt)$. In the
second, $\cos\psi\,\sin(m\psi)\sin(n\psi)$ integrates to
$\frac{\pi}{4}(\delta_{n,m+1}+\delta_{n,m-1})$, which produces exactly
the weights $(n-1)I_{n-1}^2+(n+1)I_{n+1}^2$ of $a_n$.
\end{proof}

\begin{lemma}[exact; entry law]\label{lem:entry}
$\partial_t k(0^+,\psi)=2H(\psi)$, where
$H(\psi)=\beta\sin(\psi/2)\,I_1(2\beta\cos(\psi/2))$.
\end{lemma}

\begin{proof}
Differentiate in $\psi$ the squared Graf identity
$I_0^2+2\sum_{m\ge1}I_m^2\cos(m\psi)=I_0(2\beta\cos(\psi/2))$
\cite[\S10.23(ii)]{DLMF} and compare with
$\partial_tk(0^+,\psi)=4\sum mI_m^2\sin(m\psi)$.
\end{proof}

\begin{proposition}[certified negative; no monotone full-path coupling]
\label{prop:coupling}
At $\beta=20$ the time-$3$ marginal
$\rho_t(u)\propto(Q^2s_1)(u)\,q_1(u,t)$ is not stochastically monotone in
$t$: with $a=\tfrac{4802}{10^4}$, $t_1=\tfrac{29621}{10^4}$,
$t_2=\tfrac{30070}{10^4}$, the tail masses satisfy $T(t_2,a)<T(t_1,a)$,
with certified enclosure
\begin{align*}
T(t_2,a)-T(t_1,a)\in
[-2.5299150980081786690028107,&\\
 -2.5299150980081786690028074&]\times10^{-10}.
\end{align*}
Consequently no monotone coupling of full bridge paths exists at
$\beta=20$. The certificate uses exact termwise tails
($J(m,n,a)=\int_a^\pi\sin mu\sin nu\dd u$ in closed form: no quadrature
error), positive-series Bessel enclosures with geometric tails, explicit
truncation bounds, nested passes $(M,\mathrm{prec})=(120,350)\subset
(140,500)$, and two independent interval implementations
(\texttt{mpmath.iv}; Arb via \texttt{python-flint}).
\end{proposition}

\begin{proposition}[certified negatives; the weight--kernel matrix]
\label{prop:matrix}
At $\beta=20$, for every entry weight $Q^{k-1}s_1$ ($k=1,2,3,4$) the
marginal one $q_1$-step from the moving endpoint fails stochastic
monotonicity, with certified witnesses sharing $(t_1,t_2)=(3,\,61/20)$:
$k=1$ (two-step bridge midpoint, the minimal counterexample), $a=9/10$,
margin $\approx-1.34\times10^{-2}$; $k=2$, $a=3/4$,
$\approx-1.80\times10^{-6}$; $k=4$, $a=3/10$,
$\approx-4.75\times10^{-13}$ ($k=3$ is
Proposition~\ref{prop:coupling}). Two smoothing steps repair all tested
weights (verified); one never does. The statement is relative to the
natural weight family: $k_2$ possesses its own total-positivity corner
($\beta\gtrsim8$) reachable by two-atom adversarial weights.
\end{proposition}

\begin{lemma}[exact; the $3/2$ mechanism]\label{lem:mech}
$\displaystyle
\lim_{u\downarrow0}\frac{\partial_u\partial_t\log q_1(u,t)}
{\beta\sin t\;u}=1+\frac{2\beta}{3}\cos t .$
For $t>\pi/2$ the corner sign change of $\partial_u\partial_t\log q_1$
occurs exactly when $\beta|\cos t|>3/2$ --- the same threshold as the
$TP_2$ failure of $q_1$.
\end{lemma}

\begin{proof}
$\partial_t\log q_1=-\beta\sin t\cos u+\beta\cos t\sin u\coth(\beta\sin
t\sin u)$ (direct differentiation of \eqref{eq:q1}); expand
$\coth z=z^{-1}+z/3+O(z^3)$ and differentiate in $u$; the stated limit is
a one-line computation (symbolic residual zero).
\end{proof}

\begin{remark}[verified; the mechanism reading]
The identification of this local defect as \emph{the} engine of
Propositions~\ref{prop:coupling}--\ref{prop:matrix} --- sine-type entry
weights, vanishing linearly at $0$ with mass on both sides of the
non-monotone region, flip the covariance
$\mathrm{Cov}_{\rho_t}(\mathbf 1_{u\ge a},\partial_t\log q_1)$, while the
flat weight does not --- is supported by the full weight--kernel matrix
and the $\beta$-scan of the violations (onset at $\beta\gtrsim12$,
consistent with the threshold), but is a verified reading, not a theorem.
\end{remark}

\section{The single-Bessel reduction and the master formula}
\label{sec:reduction}

\begin{lemma}[exact; Neumann product formulas]\label{lem:neumann}
For $x,y\ge0$, with $w=\sqrt{x^2+y^2+2xy\cos\alpha}$,
\[
I_0(x)I_0(y)=\frac{1}{2\pi}\int_0^{2\pi} I_0(w)\dd\alpha,
\qquad
I_0(x)I_1(y)=\frac{1}{2\pi}\int_0^{2\pi} I_1(w)\,\frac{y+x\cos\alpha}{w}
\dd\alpha .
\]
\end{lemma}

\begin{proof}
The first is the integrated Graf/Neumann addition theorem
\cite[\S10.23(ii)]{DLMF}; the second is its $\partial_y$-derivative.
\end{proof}

\begin{remark}[signs]\label{rem:signs}
Both formulas extend to all real $x,y$, as used below (on
$[-\pi,\pi]^2$ one has $c_1=\cos(\varphi/2)\ge0$ but $c_2$ may be
negative): $w(x,-y,\alpha)=w(x,y,\pi-\alpha)$, and the substitution
$\alpha\mapsto\pi-\alpha$ over a full period maps the first identity to
itself (both sides even in $y$) and the second to its negative on both
sides simultaneously ($I_1$ odd; $(-y+x\cos(\pi-\alpha))=
-(y+x\cos\alpha)$).
\end{remark}

With $K_2(\varphi)=I_0(2\beta\cos(\varphi/2))=\sum_{m\in\mathbb Z}I_m^2
e^{im\varphi}$ and $P_4=K_2*K_2$ (circle convolution),
$P_4=\sum_{m}I_m^4e^{imt}$ and $\FB=-\tfrac12P_4'$. Applying
Lemma~\ref{lem:neumann} to both convolution factors:

\begin{lemma}[exact; master representation]\label{lem:master}
Let $s=\varphi-t/2$, $c_1=\cos(\varphi/2)$, $c_2=\cos((t-\varphi)/2)$,
$P=\sin^2(s/2)$, $Q=\sin^2(\alpha/2)$, and
\begin{equation}\label{eq:R2}
R^2=c_1^2+c_2^2+2c_1c_2\cos\alpha
   =4\bigl[\cos^2(t/4)\,(1-P-Q)+PQ\bigr] .
\end{equation}
Then
\begin{equation}\label{eq:FBfull}
\FB(t)=\frac{\beta\sin(t/2)}{8\pi^2}
\iint_{[-\pi,\pi]^2}\frac{I_1(2\beta R)}{R}
\bigl[\cos^2(s/2)-\sin^2(\alpha/2)\bigr]\dd s\dd\alpha,
\end{equation}
and, under the common positive kernel $K=I_1(2\beta R)/R$,
\begin{equation}\label{eq:masterE}
E(t)=\frac{\bigl\langle\,
\cos(t/2)\cos2s+\cos\alpha\,[\cos(t/2)\cos s-\sin^2 s]\,\bigr\rangle}
{\bigl\langle\,\cos s+\cos\alpha\,\bigr\rangle},
\end{equation}
where $\langle\cdot\rangle$ denotes the $K\dd s\dd\alpha$-integral over
$[-\pi,\pi]^2$. The prefactor collapse behind \eqref{eq:FBfull} is the
sum-to-product symmetrization
$\tfrac12[\sin(\tfrac{t-\varphi}{2})(c_2+c_1\cos\alpha)+(\varphi
\leftrightarrow t-\varphi)]=\tfrac12\sin(t/2)(\cos s+\cos\alpha)$, and
$\cos s+\cos\alpha=2[\cos^2(s/2)-\sin^2(\alpha/2)]$.
\end{lemma}

\begin{proof}
(i) The squared Graf identity gives $K_2(\varphi)=\sum_mI_m^2
e^{im\varphi}$; convolution multiplies Fourier coefficients, so
$P_4=\sum_mI_m^4e^{im\varphi}$ and, termwise (super-exponential decay),
$\FB=-\tfrac12P_4'$. (ii) Insert Lemma~\ref{lem:neumann} (first formula)
with $x=2\beta c_1$, $y=2\beta c_2$ into $P_4(t)=\frac1{2\pi}\int
K_2(\varphi)K_2(t-\varphi)\dd\varphi$: a two-dimensional integral of
$I_0(2\beta R)$ with $R$ as in \eqref{eq:R2}; the $(P,Q)$ form of $R^2$
is product-to-sum. (iii) Differentiate under the integral (smooth
integrand, compact domain): $\partial_tI_0(2\beta R)=I_1(2\beta R)\cdot
2\beta\,\partial_tR$ and $2R\,\partial_tR=2(c_2+c_1\cos\alpha)\,
\partial_tc_2$ with $\partial_tc_2=-\tfrac12\sin((t-\varphi)/2)$.
(iv) Symmetrize over $\varphi\leftrightarrow t-\varphi$ ($R$ invariant):
with $\sin(x/2)\cos(x/2)=\tfrac12\sin x$ and the angle-addition formula,
\begin{align*}
\tfrac12\bigl[\sin(\tfrac{t-\varphi}{2})(c_2+c_1\cos\alpha)
+\sin(\tfrac{\varphi}{2})(c_1+c_2\cos\alpha)\bigr]
&=\tfrac14[\sin(t-\varphi)+\sin\varphi]
+\tfrac12\sin(\tfrac t2)\cos\alpha\\
&=\tfrac12\sin(\tfrac t2)(\cos s+\cos\alpha),
\end{align*}
using $\sin(t-\varphi)+\sin\varphi=2\sin(t/2)\cos s$ and
$\sin(\tfrac{t-\varphi}{2})c_1+\sin(\tfrac{\varphi}{2})c_2
=\sin(t/2)$. Collecting constants yields \eqref{eq:FBfull}. The same
substitution in the companion representation of $\FA$
(Lemma~\ref{lem:bridge} with Lemma~\ref{lem:neumann}, second formula)
and cancellation of the common factor $\beta\sin(t/2)$ yields
\eqref{eq:masterE}. As an end-to-end check, \eqref{eq:FBfull} and
\eqref{eq:masterE} match the quartic series with ratio
$1.000000000000$ at $t\in\{0.4,1.1,2.3,2.9\}$, $\beta=4$.
\end{proof}

\begin{lemma}[exact; geometry of the $(P,Q)$ square, $0<t<\pi$]
\label{lem:geometry}
On $[0,1]^2$: the main saddle is $(0,0)$ with $R=2\cos(t/4)$; the
secondary saddle is $(1,1)$ with $R=2\sin(t/4)$ and gap
$2(\cos(t/4)-\sin(t/4))>0$ degenerating only at $t=\pi$; the zero set
$\{R=0\}$ consists of exactly the two corners $(1,0)$ and $(0,1)$, and at
both of them \emph{both} brackets of \eqref{eq:FBfull} and
\eqref{eq:masterE} vanish, so the singular points of $K$ are switched off
(the extension $I_1(z)/z\to\tfrac12$ makes $K$ continuous with value
$\beta$). At $t=0$ the zero set degenerates to the edges $P=1$, $Q=1$.
The saddle deficit is the exact sum of two squares
\begin{equation}\label{eq:deficit}
4\cos^2(t/4)-R^2=4\cos^2(t/4)\sin^2(s/2)+4c_1c_2\sin^2(\alpha/2),
\qquad c_1c_2=\cos^2(t/4)-\sin^2(s/2),
\end{equation}
with sign region $c_1c_2\ge0\iff|s|\le\pi-t/2$.
\end{lemma}

\begin{proof}
\eqref{eq:R2} and \eqref{eq:deficit} are product-to-sum computations
(symbolic zeros). $R^2\ge(|c_1|-|c_2|)^2$ with equality forcing
$|c_1|=|c_2|$ (i.e.\ $s=0$ or $s=\pm\pi$) and $\cos\alpha=\mp
\mathrm{sign}(c_1c_2)$, which lands exactly on $(P,Q)\in\{(0,1),(1,0)\}$.
Bracket vanishing at both points is direct substitution. The Jacobian of
$(s,\alpha)\mapsto(P,Q)$ carries the arcsine weight
$\dd s\dd\alpha=\dd P\dd Q/\sqrt{P(1-P)Q(1-Q)}$ (per quadrant), singular
at all four corners; near the main saddle one therefore works in
$(s,\alpha)$, and in $(P,Q)$ only for the global geometry.
\end{proof}

\begin{lemma}[exact; saddle cancellations]
\label{lem:cancel}
Write $N,D$ for numerator and denominator brackets of \eqref{eq:masterE},
$C=\cos(t/2)$, $c_0=\cos(t/4)$. At the saddle $(s,\alpha)=(0,0)$:
\[
\frac{N}{D}\Big|_{(0,0)}=C,\qquad
N_{\alpha\alpha}-C\,D_{\alpha\alpha}=0,\qquad
N_{ss}-C\,D_{ss}=-4C-2,
\]
and for $N_t=-\tfrac12\sin(t/2)[\cos2s+\cos\alpha\cos s]$:
\[
N_{t,\alpha\alpha}-m\,D_{\alpha\alpha}=0\ \ (m=-\tfrac12\sin(t/2)),
\qquad N_{t,ss}-m\,D_{ss}=2\sin(t/2).
\]
Also exact: $2C+1=4c_0^2-1$ and $C\,c(t)=2c_0-1/(2c_0)$ for
$c(t)=(4c_0^2-1)/(2c_0C)$.
\end{lemma}

\begin{proposition}[verified; the second-order laws]\label{prop:secondorder}
The frozen-kernel expansion built on Lemma~\ref{lem:cancel} (with
$\langle s^2\rangle=1/(\beta c_0)+O(\beta^{-2})$ from
\eqref{eq:deficit}) predicts
\begin{equation}\label{eq:secondorder}
E=C\Bigl(1-\frac{c(t)}{\beta}\Bigr)+O(\beta^{-2}),\qquad
E'=-\tfrac12\sin(t/2)+\frac{\sin(t/4)(4c_0^2+1)}{8\beta c_0^2}
+O(\beta^{-2}).
\end{equation}
The coefficient identities are exact (Lemma~\ref{lem:cancel}); the
\emph{laws with remainder} are verified, not yet certified: measured
remainders are $\approx0.15/\beta^2$ with true convergence at fixed $t$
(e.g.\ $\beta^2$-scaled residues $0.1776/0.1767/0.1762$ at $t=1$,
$\beta=30/60/120$), and $\beta\,(E'+\tfrac12\sin(t/2))\to
\sin(t/4)(4c_0^2+1)/(8c_0^2)$ at $t=1,2$. The remainder control is
exactly what the closure map owes. Note the $1/\beta$ correction to $E'$
is \emph{positive}: it opposes the target inequality and must be beaten,
not used. The $\alpha$-cancellations imply every kernel correction
multiplies a vanishing fluctuation and enters at $O(\beta^{-2})$ only.
\end{proposition}

\begin{remark}[exact decomposition of $E'$]\label{rem:Eprime}
With the normalized positive measure $\dd\mu\propto K\dd s\dd\alpha$,
$E'=\langle N_t\rangle/\langle D\rangle+\langle(N-ED)\,
\partial_t\log K\rangle/\langle D\rangle$, $D_t=0$, and the regular form
$\partial_t\log K=4cc'(1-P-Q)\,[zI_0(z)/I_1(z)-2]/R^2$, $z=2\beta R$,
extends continuously by $4\beta^2cc'(1-P-Q)$ at $R=0$. Verified split at
$(\beta,t)=(5,1.3)$: $-0.238212-0.024552=-0.262764=$ the finite
difference exactly. Bounds for $E'$ are taken from this exact quotient
--- never by differentiating an estimate of $E$.
\end{remark}

\section{The $\pi$-endpoint package}\label{sec:endpoint}

\begin{lemma}[exact; telescoping]\label{lem:telescope}
With $a_m$ as in \eqref{eq:coeffs}:
\[
A_\infty:=\sum_{m\ge1}(-1)^{m+1}m\,a_m=0,
\qquad
\sum_{m\ge1}(-1)^{m+1}m^3a_m
=-\sum_{m\ge1}(-1)^{m+1}m(m+1)(2m+1)\,I_m^2I_{m+1}^2 .
\]
Hence $\FA(\pi-\varepsilon)=c_3\,\varepsilon^3+O(\varepsilon^5)$ with
\begin{equation}\label{eq:c3}
c_3=\frac16\sum_{m\ge1}(-1)^{m+1}m(m+1)(2m+1)\,I_m^2I_{m+1}^2 .
\end{equation}
\end{lemma}

\begin{proof}
Split $m^ra_m=m^r(m-1)I_{m-1}^2I_m^2+m^r(m+1)I_m^2I_{m+1}^2$ and shift
$m\to m+1$ in the first sum: for $r=1$ the two weights cancel
($m(m+1)-m(m+1)=0$); for $r=3$ they combine to
$m(m+1)[m^2-(m+1)^2]=-m(m+1)(2m+1)$. Absolute convergence justifies the
reindexing; $\sin(m(\pi-\varepsilon))=(-1)^{m+1}\sin(m\varepsilon)$
expands the sine.
\end{proof}

\begin{lemma}[exact; integral form and parity]\label{lem:c3int}
Let $g(u)=\tfrac12 I_1(2\beta\cos u)=\sum_{m\ge0}I_mI_{m+1}
\cos((2m+1)u)$ \textup{(addition theorem)}. Then
\[
c_3=\frac{S_3-S_1}{24},\qquad
S_1=\frac{2}{\pi}\int_0^{\pi}g(\tfrac\pi2-u)\,g'(u)\dd u,\quad
S_3=\frac{2}{\pi}\int_0^{\pi}g'(\tfrac\pi2-u)\,\bigl(-g''(u)\bigr)\dd u .
\]
Moreover $g(-u)=g(u)$ and $g(\pi-u)=-g(u)$ (parity of $I_1$), hence
$g'(\pi-u)=g'(u)$, $g''(\pi-u)=-g''(u)$, and \emph{both} integrands are
invariant under $u\mapsto\pi-u$: the two halves $[0,\pi/2]$ and
$[\pi/2,\pi]$ contribute exactly equally, and one may integrate a half
range and double. In particular no cancellation between the two phase
maxima $u=\pi/4$ and $u=3\pi/4$ is possible --- exactly, not
asymptotically.
\end{lemma}

\begin{proof}
The expansion of $g$ is the Graf addition theorem with unit shift.
Quarter turn: $\cos((2m+1)(\tfrac\pi2-u))=(-1)^m\sin((2m+1)u)$, so
$g(\tfrac\pi2-u)=\sum_m(-1)^mI_mI_{m+1}\sin((2m+1)u)$ while
$g'(u)=-\sum_k(2k+1)I_kI_{k+1}\sin((2k+1)u)$ and
$g''(u)=-\sum_k(2k+1)^2I_kI_{k+1}\cos((2k+1)u)$, whence
$g'(\tfrac\pi2-u)=-\sum_k(-1)^k(2k+1)I_kI_{k+1}\cos((2k+1)u)$.
Orthogonality of $\{\sin((2m+1)u)\}$ and $\{\cos((2m+1)u)\}$ on
$[0,\pi]$ (norm $\pi/2$) gives
$S_1=-\sum_m(-1)^m(2m+1)(I_mI_{m+1})^2$ and
$S_3=-\sum_m(-1)^m(2m+1)^3(I_mI_{m+1})^2$, so
\begin{align*}
\frac{S_3-S_1}{24}
&=\sum_{m\ge0}(-1)^{m+1}\frac{(2m+1)^3-(2m+1)}{24}(I_mI_{m+1})^2\\
&=\frac16\sum_{m\ge1}(-1)^{m+1}m(m+1)(2m+1)I_m^2I_{m+1}^2=c_3,
\end{align*}
using $(2m+1)^3-(2m+1)=4m(m+1)(2m+1)$ and the vanishing of the $m=0$
term. The parity claims follow from $I_1(-z)=-I_1(z)$ applied to
$g(\pi-u)=\tfrac12I_1(-2\beta\cos u)$, and by differentiating the two
symmetry relations.
\end{proof}

\begin{proposition}[verified; prefactor law]\label{prop:prefactor}
$c_3>0$ for $\beta\in\{1,5,20,40,80,120\}$ at cancellation-adjusted
precision \textup{(the alternating sum cancels at scale
$e^{-(4-2\sqrt2)\beta}$; evaluations use $\mathrm{dps}\ge0.6\beta+50$)},
and
\[
c_3\;\sim\;A\,\beta^{3/2}e^{2\sqrt2\beta},\qquad
A=\frac{1}{24\cdot2^{1/4}\pi^{3/2}}=0.0062922\ldots,
\]
with measured ratios $0.0058646,\,0.0060049,\,0.0060996,\,0.0061475$ at
$\beta=40,60,90,120$ and $O(1/\beta)$ gaps. The certified large-$\beta$
branch (pending) must produce the signed prefactor with an explicit
$B$: $I_{\rm half}\ge\frac{A}{2}\beta^{3/2}e^{2\sqrt2\beta}(1-B/\beta)
-T(\beta)$, $c_3=2I_{\rm half}$, read $\beta^*$, and close
$[15,B_{\rm fin}]$, $B_{\rm fin}=\max\{\beta^*,10C\}$, by intervals.
\end{proposition}

\section{The closure map and the candidate threshold}\label{sec:closure}

\begin{proposition}[verified truth map; all statements on the stated
finite grids]\label{prop:truthmap}
Let $g(t,\beta)=E'+\tfrac14\sin(t/2)$. At $\beta=15$, $g<0$ at every
point of a $13$-point grid spanning $[0.05,\pi-0.09]$ (independently
reproduced on a much finer mesh at $110$ digits), with thinnest margin $-0.00573$ at $t=0.05$ (on the
seal window $[0.05,\pi-0.10]$) and $-0.00216$ at $\pi-0.09$; the
crossing sits at $\pi-t^*=0.0890502225$. At $\beta=12$ the truth itself
fails at $\pi-0.10$ ($g=+0.025$); crossings scale as the mirror window:
$\pi-t^*=0.1155,\,0.0891,\,0.0644$ at $\beta=12,15,20$, matching
$\sqrt2/\beta$. A sweep $\beta=15\ldots60$ keeps the maximum at $t=0.05$,
drifting to $-0.0061$: strong uniformity evidence, not proof.
\end{proposition}

\begin{conjecture}[the pending rows of the relay, quantified]\label{conj:seal}
Part \textup{(ii)} of Conjecture~\ref{conj:main} holds with the two-scale
decomposition: \textup{(a)} $0<\beta\le3$ by Theorem~\ref{thm:B};
\textup{(b)} $\beta\in[3,15]$ by certified evaluation \textup{(the
holonomic structure of Section~\ref{sec:verif} makes the compact cheap)};
\textup{(c)} $\beta\ge15$ by three uniform blocks \textup{(windows per
the LIVE map of Section~\ref{sec:relay}, superseding earlier drafts)}:
left edge $(0,0.6]$
\textup{(}$e_2>0$, proved by the Chebyshev-sum corollary of the
Tur\'an lemma, plus a remainder radius $0.6$ via the $e_4$
coefficient\textup{)}; bulk
$[0.6,\pi-C/\beta]$ \textup{(}enclosure of $\langle s^2\rangle$ and of the
covariance of Remark~\ref{rem:Eprime}, with the five
prototype-to-certificate conversions\textup{)}; right edge with
\emph{moving} boundary: assembly to $\pi-C/\beta$, then the quadratic
regime of $c_3$ on $[\pi-C/\beta,\pi]$ with $C_{\rm mirror}\le C\le
C_{\rm quad}$ \textup{(}a fixed quadratic window is provably impossible:
$\kappa\approx0.11\beta$ grows while $E(\pi-0.10)$ stays bounded;
retention $E/\kappa d^2$ at $d=2/\beta$ is $0.896,0.892,0.890$ at
$\beta=20,40,80$ --- the validity radius scales as $1/\beta$\textup{)}.
The named constants still owed: $B$, $\beta^*$, $C_{\rm quad}$,
$C_{\rm mirror}$, and the remainder function $T(\beta)$.
\end{conjecture}

The first-order competition is not the obstacle: the two-term law
\eqref{eq:secondorder} requires only
$\beta>(1+1/(4c_0^2))/c_0\le3\sqrt2/2\approx2.121$; everything above that
is rigor slack. The candidate $\beta_0=15$ is near truth-optimal for its
edge window (Proposition~\ref{prop:truthmap}), not padded.

\section{Verification, certificates, and reproducibility}\label{sec:verif}

All numerical artifacts live in the repository
\begin{center}
\url{https://github.com/lluiseriksson/THE-ERIKSSON-PROGRAMME}
\end{center}
(directories \texttt{scripts/}, \texttt{docs/}); the immutable commit
hash for the version audited with this manuscript is recorded in the
repository's hash registry (\texttt{docs/SURFACE-CLOSURE-NOTES.md}) and
in the arXiv comment field on submission. Each certificate embeds its
pinned statement and self-verifies on every execution.

\begin{sloppypar}
\emph{Certificates (interval arithmetic, outward rounding, nested or
stability passes).} \texttt{certify\_time3\_negative.py} and its Arb twin
(Proposition~\ref{prop:coupling}); \texttt{certify\_bridge\_matrix.py}
and twin (Proposition~\ref{prop:matrix}); \texttt{certify\_thmB.py}
(Theorem~\ref{thm:B}: adaptive $\beta$-boxes covering $[1/20,3]$;
tails computed in intervals and closed at the level of
\emph{complete shells}: each pair maps into the next shell with term
ratio $r$ derived from the termwise-exact
$I_{n+1}(x)\le xI_n(x)/(2(n+1))$, and the one new pair per shell is
bounded by a double index shift, giving shell ratio $2r$, asserted
$<\tfrac12$ per box before the geometric closure; a built-in full stability pass at
precision $+70$; the Arb twin \texttt{certify\_thmB\_arb.py} independently
certifies the full coverage --- $86$ boxes, both passes, transcript
archived; the \texttt{mpmath.iv} canonical transcript is committed on
completion of its run, as recorded in the repository hash registry). Bessel enclosures use the positive
series with an exact-integer stopping rule and geometric tails; floats
appear only in loop-termination heuristics. Nested enclosures prevent
silent numerical or truncation degradation; statement auditing and the
independent Arb implementation guard against common-mode errors.
\end{sloppypar}

\emph{Verified computations.} The master-formula and reduction identities
(ratios $1.0$ to $12$ digits); the truth map of
Proposition~\ref{prop:truthmap} (dps $80$, confirmed independently at
$110$ digits); the $c_3$ ladder (Proposition~\ref{prop:prefactor}); the
term-split coefficients of \eqref{eq:secondorder} at
$\beta=60,120,240$.

\emph{Precision policies} (each traceable to a caught artifact):
$\mathrm{dps}\ge2.2\beta+40$ for any evaluation with $t\ge2$, $\beta>30$
(mirror cancellation); $\mathrm{dps}\ge0.6\beta+50$ for $c_3$-type
alternating sums; series cutoffs scale with the maximal argument, and
scaled-asymptotic evaluation of $e^{-z}I_n(z)$ is the default for
$z>35$; every measured coefficient carries a $\beta$-scaling sanity
check; whatever can be computed from a positive integral representation
is.

\emph{Audit trail.} Fifteen numerical or formulation artifacts
(``ghosts'') were caught during fabrication and writing by the three-way
review protocol; Appendix~\ref{app:ghosts} tabulates them. We consider
the table part of the result: it is the reproducibility record that
makes the certificates credible.

\begin{sloppypar}
\emph{Machine-checked lemmas.} The unit-step and Tur\'an-type coefficient
inequalities behind Theorem~\ref{thm:B} are formalized in Lean~4/Mathlib
(\texttt{PhiMonotone.lean}, \texttt{FHBesselAmos.lean},
\texttt{WronskianIdentity.lean}; axiom oracle
\texttt{[propext, Classical.choice, Quot.sound]}, no \texttt{sorry}).
Mathlib has no Bessel functions: what is machine-checked is the
ordered-field implication, with the analytic inputs (three-term
recurrence \cite[10.29.1]{DLMF}; the Amos-type bound of
\cite{Amos,Segura2011,RuizSegura}, cited verbatim, shift
$\nu\to\nu+1$ as in Appendix~\ref{app:phi}) carried as named
hypotheses. Correct claim everywhere: \emph{machine-checked modulo
classical Bessel inputs carried as named hypotheses.}
\end{sloppypar}

\section*{Acknowledgements and provenance}
The conjecture, the bridge reading, the reduction, and all certificates
were produced inside an audit-first programme in which three independent
reviewing voices verify every claim before acceptance; scores,
objections, and the complete ghost ledger are archived with the
repository history. Nothing in this paper rests on trust in any single
computation: every load-bearing negative has two witnesses, and every
identity marked exact has a symbolic-zero transcript.

\appendix

\section{The weighted Tur\'an-type lemma, self-contained}\label{app:phi}

\begin{theorem}[exact]\label{thm:phi}
For every $x>0$ and every real $m\ge1$, with $\varphi_m=a_m/b_m$ built
from \eqref{eq:coeffs} at argument $x$:\; $\varphi_m<\varphi_{m+1}$.
Equivalently, with $\rho_\mu=I_{\mu+1}/I_\mu$,
\[
\frac1m\Bigl(\frac{m-1}{\rho_{m-1}^2}+(m+1)\rho_m^2\Bigr)
<\frac1{m+1}\Bigl(\frac{m}{\rho_m^2}+(m+2)\rho_{m+1}^2\Bigr).
\]
\end{theorem}

\begin{proof}
Two classical inputs, valid for all real orders: the three-term
recurrence $\rho_{m-1}^{-1}=\rho_m+2m/x$, $\rho_{m+1}=\rho_m^{-1}
-2(m+1)/x$ \cite[10.29.1]{DLMF}, and the Amos-type bound
\[
\rho_m(x)<U_m(x):=\frac{x}{a+\sqrt{a^2+x^2}},\qquad a=m+\tfrac12,
\]
proved best of its form for all real orders $\ge0$ by Segura
\cite{Segura2011}; we use \cite[Thm.~2]{RuizSegura}, which bounds
$I_\nu(x)/I_{\nu-1}(x)$ for $\nu\ge\tfrac12$ by
$x/\bigl(\nu-\tfrac12+\sqrt{(\nu+\tfrac12)^2+x^2}\,\bigr)$; setting
$\nu=m+1$ bounds $\rho_m=I_{m+1}/I_m$ by
$x/\bigl(m+\tfrac12+\sqrt{(m+\tfrac32)^2+x^2}\,\bigr)
\le x/\bigl(a+\sqrt{a^2+x^2}\,\bigr)$ with $a=m+\tfrac12$, which is the
displayed bound (the last step uses $\sqrt{(m+\tfrac32)^2+x^2}\ge
\sqrt{a^2+x^2}$).
Substitute the recurrences, write $u=\rho_m\in(0,1)$, $c=1/x$,
$S=u+1/u$, $P=1/u-u$; clearing the positive denominators, the claim
becomes $\Delta_m>0$ where the \emph{exact factorization}
\[
\frac{\Delta_m}{2m(m+1)}=(S-3c)\bigl(P-(2m+1)c\bigr)+(2m+1)c^2
\]
holds as an algebraic identity (expand and collect; the coefficients of
$u^{\pm2}$ are $\pm2m(m+1)$, the linear-in-$c$ terms are
$-4m(m+1)c[(m-1)u+(m+2)/u]$, and $(m-1)u+(m+2)/u=\tfrac12[(2m+1)S+3P]$,
$1/u^2-u^2=PS$). The second factor is positive because $U_m$ is the
positive root of $1/w-w=(2m+1)/x$ (the calibration $1/U_m-U_m=2a/x$) and
$w\mapsto1/w-w$ is strictly decreasing, so $\rho_m<U_m$ gives
$P>(2m+1)c$. The first factor is positive because $\sqrt{a^2+x^2}>a$
gives $U_m<x/(2a)$, so $u<x/(2m+1)\le x/3$ and
$S>1/u>(2m+1)/x\ge3c$. The third term is positive. Strictness is strict
throughout. \textup{(}This is the proof of \cite{PhiLemma}, reproduced
so that Theorem~\ref{thm:B} is self-contained; the ordered-field
implication is machine-checked in \texttt{PhiMonotone.lean}, modulo the
two classical inputs carried as named hypotheses.\textup{)}
\end{proof}

\section{Audit table (the ghost ledger)}\label{app:ghosts}

Each row: what was wrong, how it was caught, and the rule it produced.
All entries are documented with transcripts in
\texttt{docs/SURFACE-CLOSURE-NOTES.md}.

\begin{center}\small
\begin{tabular}{@{}rp{5.0cm}p{7.6cm}@{}}
\toprule
\# & artifact & catch / rule produced\\
\midrule
1 & random $TP_2$ scan missed adversarial corners & targeted corner
search; adversarial grids mandatory\\
2 & verification cell with unbound parameter & explicit binding rule for
inline prints\\
3 & near-$\pi$ scan at low precision, false violation & precision rule
$\mathrm{dps}\ge2.2\beta+40$ for $t\ge2$, $\beta>30$\\
4 & stale-PDF claim misattributed between voices & provenance ledger:
attribute to exact source\\
5 & float64 CDF increments $\pm11$ (impossible) & positive-representation
preference\\
6 & $t{=}2$, $\beta{=}240$ residue $0.363$ vs $0.317$ & boundary $t\ge2$
included in the dps rule\\
7 & $I_1$ series truncated at $60$ terms, $z\le480$ & cutoffs scale with
argument; scaled asymptotics for $z>35$; $\beta$-scaling checks\\
8 & $c_3$ sign flip at $\beta=120$ (cancellation) & alternating-sum rule
$\mathrm{dps}\ge0.6\beta+50$\\
9 & $R=0$ corner misplaced at $(1,1)$ & complete critical-point census
required in prose too\\
10 & one-saddle prefactor off by exactly $2$ & census rule; the factor
$2$ later proved an exact parity\\
11 & $I_1$ lower bound dropped the negative tail & sign audits are a
separate mandatory pass\\
12 & fixed quadratic window at $\pi$ was provably false & interrogate
obligations before attempting them\\
13 & Thm.~A second proof: rearrangement misquoted & layer-cake arc proof
substituted (this revision)\\
14 & Lemma sign error: floor stated as upper bound & corrected; per-pair
statements now carry signs explicitly\\
15 & ``exact'' labels on verified inequalities & tricotomy re-audit of
every label (this revision)\\
\bottomrule
\end{tabular}
\end{center}

\begin{thebibliography}{99}
\bibitem{Amos} D.~E.~Amos, \emph{Computation of modified Bessel functions
and their ratios}, Math.\ Comp.\ \textbf{28} (1974), 239--251.
\bibitem{Segura2011} J.~Segura, \emph{Bounds for ratios of modified
Bessel functions and associated Tur\'an-type inequalities}, J.\ Math.\
Anal.\ Appl.\ \textbf{374} (2011), 516--528.
\bibitem{RuizSegura} D.~Ruiz-Antol\'in, J.~Segura, \emph{A new type of
sharp bounds for ratios of modified Bessel functions}, J.\ Math.\ Anal.\
Appl.\ \textbf{443} (2016), 1232--1246.
\bibitem{DLMF} NIST Digital Library of Mathematical Functions,
\url{https://dlmf.nist.gov}.
\bibitem{Companion} L.~Eriksson, \emph{Recurrence--Amos proof of the
unit-step order-monotonicity of $(\log I_\nu)'$}, ai.viXra:2607.0020
(2026).
\bibitem{PhiLemma} L.~Eriksson, \emph{A weighted Tur\'an-type
monotonicity lemma for modified Bessel functions via the calibrated Amos
bound}, ai.viXra:2607.0017 (2026).
\bibitem{WronskianNote} L.~Eriksson, \emph{A Wronskian identity for
antisymmetrized sine sums}, ai.viXra (2026), number pending.
\bibitem{ParityNote} L.~Eriksson, \emph{Parity barriers for decoupling
inequalities}, ai.viXra:2607.0018 (2026).
\bibitem{Watson} G.~N.~Watson, \emph{A Treatise on the Theory of Bessel
Functions}, 2nd ed., Cambridge Univ.\ Press, 1944.
\bibitem{KarlinMcGregor} S.~Karlin, J.~McGregor, \emph{Coincidence
probabilities}, Pacific J.\ Math.\ \textbf{9} (1959), 1141--1164.
\bibitem{MardiaJupp} K.~V.~Mardia, P.~E.~Jupp, \emph{Directional
Statistics}, Wiley, 2000.
\bibitem{DJD} S.~Dharmadhikari, K.~Joag-Dev, \emph{Unimodality,
Convexity, and Applications}, Academic Press, 1988.
\end{thebibliography}

\end{document}
